---
title: "Founders Intravenous Self-Administration Data: Heritability and Plotting"
author: "Michael C. Saul (michael.saul [at] jax.org)"
modified by: "Udita Datta (udita.datta [at] jax.org)"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: true
    fig_retina: true
    self_contained: true
    code_folding: show
    theme: flatly
    highlight: default
    df_print: paged
---

# Background

## Motivation

### Purpose of analysis

This analysis is performed to process the intravenous self-administration data and calculate:

1) Heritability
2) Strain-by-sex means and standard errors for plotting

**These data were collected by the CSNA research assistants. Ultimate authorship of the resultant publications will be determined in discussion with Leona Gagnon.** The preprocessing script `IVSA_import_preprocess.Rmd` was used to filter the data to only those data requiring reporting.

### Note about this analysis

The data have a floor effect (animals that took no cocaine) for infusions and a ceiling effect (animals that did not acquire self-administration) for sessions to acquisition. The appropriateness of a conventional linear model should be discussed with Vivek Philip. It's possible that we should fit a Tobit regression.

# Analysis

## Setup

### Working directory

All chunks will run in the working folder for this experiment.

```{r setup}
script_root_dir = "C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA"
knitr::opts_knit$set(root.dir = script_root_dir)
```

Getting current date

```{r}
current_date = format(Sys.time(),"%Y-%m-%d")
```

### R libraries

Calling R libraries necessary for this analysis.

```{r}
library("tidyverse")
library("lubridate")
library("cowplot")
library("MASS")
library("ggbeeswarm")
```

## Importing data

## Importing and displaying raw intravenous self-administration data

Getting data from the filtering script and taking out the `DE_` part of the header

```{r}
IVSA_data = readRDS(paste("C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA/csna_project_4_behavior/data/IVSA_data_preprocessed_EX_RE_2020-12-20.RDS", sep = ""))
colnames(IVSA_data) = gsub("^DE_","",colnames(IVSA_data))
row.names(IVSA_data) = as.character(IVSA_data$Subject)
IVSA_data
```

Getting founders strain key

```{r}
founders_key = readRDS("C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA/csna_project_4_behavior/data/cc_do_founders_key.RDS")
row.names(founders_key) = founders_key$strain
IVSA_data = IVSA_data %>%
  filter(Strain %in% founders_key$strain)
#rm(list = c("IVSA_raw"))
IVSA_data = as.data.frame(IVSA_data)
```
### Fitting ANOVA model for Slope_Extinction_Active_Lever: Extinction phase of self-administraion

Fitting the model with genetic background (strain), sex, and strain x sex interactions.

```{r}
anova(lm(Slope_Extinction_Active_Lever ~ Strain * Sex, data = IVSA_data))
```

No effects are significant  in this dataset. Plotting residuals of model with strain, anyways.



```{r}
founders_key_no_129 <- founders_key[which(founders_key$strain != "129S1/SvImJ"),]
IVSA_resid = IVSA_data %>%
  filter(!is.na(Slope_Extinction_Active_Lever))
IVSA_resid$DR_Inf_Total_AUC_Resid = resid(lm(Slope_Extinction_Active_Lever ~ Strain, data = IVSA_resid))

ggplot(data = IVSA_resid, aes(x = Strain, y = Slope_Extinction_Active_Lever, color = Strain, shape = Sex)) +
  geom_hline(yintercept = 0, col = "#999999") +
  geom_beeswarm(size = 2) + 
  scale_color_manual(values = founders_key_no_129$collaborative_cross_color_broman) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(16,23))
```

```{r}
min(IVSA_resid$Slope_Extinction_Active_Lever)
boxcox(lm((Slope_Extinction_Active_Lever + 74) ~ Strain, data = IVSA_resid))
```

The log-likelihood centered at 0 implies that a log-transform may help with the non-normal residuals.

```{r}
IVSA_resid$Slope_Extinction_Active_Lever_Resid_constant = resid(lm((Slope_Extinction_Active_Lever + 74) ~ Strain, data = IVSA_resid))

ggplot(data = IVSA_resid, aes(x = Strain , y = Slope_Extinction_Active_Lever_Resid_constant, color = Strain, shape = Sex)) +
  geom_hline(yintercept = 0, col = "#999999") +
  geom_beeswarm(size = 2) + 
  scale_color_manual(values = founders_key_no_129$collaborative_cross_color_broman) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(16,23))
```

This looks a little better. Trying Box-Cox to see if the log-likelihood is centered at 1 with the transformed data.



```{r}
mean_n = nrow(IVSA_data) / length(unique(IVSA_data$Strain))
anova_Slope_Extinction_Active_Lever = anova(lm((Slope_Extinction_Active_Lever+74) ~ Strain, data = IVSA_data))
anova_Slope_Extinction_Active_Lever_df = as.data.frame(anova_Slope_Extinction_Active_Lever)
h2_Slope_Extinction_Active_Lever = anova_Slope_Extinction_Active_Lever_df["Strain","Mean Sq"] / (anova_Slope_Extinction_Active_Lever_df["Strain","Mean Sq"] + ((mean_n - 1) * anova_Slope_Extinction_Active_Lever_df["Residuals","Mean Sq"]))
h2_Slope_Extinction_Active_Lever
```


```{r}
mean_n = nrow(IVSA_data) / length(unique(IVSA_data$Strain))
anova_Slope_Extinction_Active_Lever = anova(lm(((Slope_Extinction_Active_Lever)^ (1/3)) ~ Strain, data = IVSA_data))
anova_Slope_Extinction_Active_Lever_df = as.data.frame(anova_Slope_Extinction_Active_Lever)
h2_Slope_Extinction_Active_Lever = anova_Slope_Extinction_Active_Lever_df["Strain","Mean Sq"] / (anova_Slope_Extinction_Active_Lever_df["Strain","Mean Sq"] + ((mean_n - 1) * anova_Slope_Extinction_Active_Lever_df["Residuals","Mean Sq"]))
h2_Slope_Extinction_Active_Lever
```
```{r}
anova(lm(RI_percent_final_EX ~ Strain * Sex, data = IVSA_data))
```

Again, strain is the only significant factor in this dataset. Plotting residuals of model with only this factor.

Strain is the only significant factor here. Looking at the residuals of the simplest model with just this in it first.

```{r}
founders_key_no_129 <- founders_key[which(founders_key$strain != "129S1/SvImJ"),]
IVSA_resid = IVSA_data %>% 
  filter(!is.na(RI_percent_final_EX))
IVSA_resid$RI_percent_final_EX_Resid = resid(lm(RI_percent_final_EX ~ Strain, data = IVSA_resid))

ggplot(data = IVSA_resid, aes(x = Strain, y = RI_percent_final_EX, color = Strain, shape = Sex)) +
  geom_hline(yintercept = 500, col = "#999999") +
  geom_beeswarm(size = 2) + 
  scale_color_manual(values = founders_key_no_129$collaborative_cross_color_broman) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(16,23))
```

This fit doesn't look good. Trying Box-Cox to see if a response variable transform may be merited.

```{r}
boxcox(lm(RI_percent_final_EX ~ Strain, data = IVSA_resid))
```


```{r}
mean_n = nrow(IVSA_data) / length(unique(IVSA_data$Strain))
anova_RI_percent_final_EX = anova(lm(((RI_percent_final_EX)^ (1/3)) ~ Strain, data = IVSA_data))
anova_RI_percent_final_EX_df = as.data.frame(anova_RI_percent_final_EX)
h2_RI_percent_final_EX = anova_RI_percent_final_EX_df["Strain","Mean Sq"] / (anova_RI_percent_final_EX_df["Strain","Mean Sq"] + ((mean_n - 1) * anova_RI_percent_final_EX_df["Residuals","Mean Sq"]))
h2_RI_percent_final_EX
```
## Document Control

This document was prepared using RMarkdown in RStudio.
