---
title: "IVSA Data: Preprocessing"
author: "Michael C. Saul (michael.saul [at] jax.org)"
adapted by: "Udita Datta (udita.datta [at] jax.org)"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: true
    fig_retina: true
    self_contained: true
    code_folding: show
    theme: flatly
    highlight: default
    df_print: paged
---

# Background

## Motivation

### Purpose of analysis

This analysis is performed to preprocess the intravenous self-administration data to:

1) Make sure all data are clean and deduplicated.
2) Make short and simple column names.
3) Clean up duplicate IDs in novelty data in a rational way.
4) Prepare data for cross-platform compatibility in both Python and R.

**These data were collected by RAs from the CSNA's behavioral core. Intellectual contributions will be assessed in discussions with Leona Gagnon.**

# Analysis

## Setup

### Working directory

All chunks will run in the working folder for this experiment.

```{r setup}
script_root_dir = "C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA"
knitr::opts_knit$set(root.dir = script_root_dir)
```

Getting current date

```{r}
current_date = format(Sys.time(),"%Y-%m-%d")
```

### R libraries

Calling R libraries necessary for this analysis.

```{r}
library("dplyr")
library("tidyverse")
library("lubridate")
library("readxl")
library("nlme")
library("ez")
```

## Importing data

## Importing IVSA and LIMS data

Getting data from the following path:

* `~/Box Sync/IVSA/preprocessed_IVSA_data/`
I have changed this to define the file path for navigating the most recent datafile on my local data folder.

```{r import_data, warning=FALSE}
# Taking the most recent datasets
data_path = paste(script_root_dir,"/preprocessed_IVSA_data/", sep = "")
IVSA_datasets = data.frame(files = list.files(path = data_path, pattern = "\\d{4}-\\d{2}-\\d{2}\\.RDS$"),
                                        date = ymd(gsub("^.*_(\\d{4}-\\d{2}-\\d{2})\\.RDS$","\\1", 
                                                        list.files(path = data_path, 
                                                                   pattern = "\\d{4}-\\d{2}-\\d{2}\\.RDS$")), 
                                                   tz = "America/New_York"),
                                        stringsAsFactors = FALSE)
most_recent_datasets = paste(data_path, IVSA_datasets[which(IVSA_datasets$date == max(IVSA_datasets$date)),"files"], sep = "")

# IVSA
IVSA_raw_file = most_recent_datasets[grep("IVSA_raw", most_recent_datasets)]
IVSA_metadata_file = most_recent_datasets[grep("IVSA_metadata", most_recent_datasets)]
IVSA_raw = readRDS(IVSA_raw_file)
row.names(IVSA_raw) = IVSA_raw$DE_Subject
IVSA_metadata = readRDS(IVSA_metadata_file)
head(IVSA_raw)
```

Getting founders strain key

```{r}
# Founders Key
founders_key = readRDS("C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA/csna_project_4_behavior/data/cc_do_founders_key.RDS")
row.names(founders_key) = founders_key$strain
founders_key[9,1] = "J:DO"
founders_key[9,2] = "DO"
founders_key[9,5] = "#000000"
founders_key[9,6] = "#000000"
row.names(founders_key) = founders_key$strain
head(IVSA_raw)
```
```{r filter_for_founders, warning=FALSE}
founders_IVSA = IVSA_raw %>%
  filter(DE_Strain %in% founders_key$strain)
#rm(list = c("IVSA_raw"))
founders_IVSA = as.data.frame(founders_IVSA)

founders_IVSA$DE_Strain = factor(founders_IVSA$DE_Strain,
                                 levels = founders_key$strain,
                                 ordered = TRUE)


Old_DO <- founders_IVSA %>% 
  filter(DE_birth_date_year == 2016,
         DE_Strain == "J:DO")


New_DO <- founders_IVSA %>% 
  filter(DE_birth_date_year != 2016,
         DE_Strain == "J:DO")

No_DO <- founders_IVSA %>% 
  filter(DE_Strain != "J:DO")

founders_IVSA <- merge(Old_DO,No_DO, all = TRUE)

ids <- c("A2063", "A2810", "A2418", "A3262", "A3261", "A3052", "A3257", "A3256", "A3051", "A2986")
rm.ids <- which(founders_IVSA$DE_Subject %in% ids)
founders_IVSA.p1  <- founders_IVSA[-rm.ids,]

founders_IVSA = founders_IVSA.p1

```
Getting CC key.

```{r}
CC_key = unique(IVSA_raw$DE_Strain)
CC_key = CC_key[grep("^CC\\d{3}/",CC_key)]
#CC_key = stringr::str_extract(CC_key, "^.{5}")
CC_key
CC_founders_IVSA_CC = IVSA_raw %>%
  filter(DE_Strain %in% c(CC_key))

CC_founders_IVSA_CC$DE_Strain = stringr::str_extract(CC_founders_IVSA_CC$DE_Strain, "^.{5}")
head(CC_founders_IVSA_CC)
```

Merging old DO, founders and CC
```{r}
CC_founders_IVSA <- rbind(founders_IVSA, CC_founders_IVSA_CC)
```


### Cleaning up IVSA data

Cleaning up IVSA data. First, selecting only founders from the LIMS file and filtering the IVSA data for these data.

## Acquisition

Excluding animals who were found dead who didn't acquire and changing the failed acquisition animals to 28 sessions.


```{r}
CC_founders_IVSA_filtered = CC_founders_IVSA %>%
  dplyr::filter(!(DE_Exit_Reason == "FD" & is.na(AQ_SessionsToAcquisition)))

CC_founders_IVSA_filtered[which(CC_founders_IVSA_filtered$DE_Exit_Reason == "IVSA Acquisition Only"),"AQ_SessionsToAcquisition"] = 28
CC_founders_IVSA_filtered
```
Looking at the rest of the data.

```{r}
CC_founders_IVSA_filtered[which(is.na(CC_founders_IVSA_filtered$AQ_SessionsToAcquisition)),]

subset(CC_founders_IVSA_filtered,is.na(CC_founders_IVSA_filtered$AQ_SessionsToAcquisition) & CC_founders_IVSA_filtered$DE_Exit_Reason == "Finished Pipeline")

```


7 individuals  appears to have finished the pipeline, but the script could not calculate a sessions to acquisition criterion for it. Excluding these data.

```{r}
CC_founders_IVSA_filtered %>%
  filter(!is.na(AQ_SessionsToAcquisition))

CC_founders_IVSA_filtered
```
Selecting the relevant columns for Extinction relevant trait calculations. Choosing the following variables:
'EX_ALP_Total_s01'to 'EX_ALP_Total_s10': Active lever press for extinction sessions
'EX_ILP_Total_s01'to 'EX_ILP_Total_s10': Inactive lever press for extinction sessions
Changing the format from wide to long format.
```{r extinction_plot, warning=FALSE}
CC_founders_extinction_df = CC_founders_IVSA %>%
  select(DE_Subject,
         DE_Strain,
         DE_Sex,
         EX_ALP_Total_s01,
         EX_ALP_Total_s02,
         EX_ALP_Total_s03,
         EX_ALP_Total_s04,
         EX_ALP_Total_s05,
         EX_ALP_Total_s06,
         EX_ALP_Total_s07,
         EX_ALP_Total_s08,
         EX_ALP_Total_s09,
         EX_ALP_Total_s10,
         EX_ILP_Total_s01,
         EX_ILP_Total_s02,
         EX_ILP_Total_s03,
         EX_ILP_Total_s04,
         EX_ILP_Total_s05,
         EX_ILP_Total_s06,
         EX_ILP_Total_s07,
         EX_ILP_Total_s08,
         EX_ILP_Total_s09,
         EX_ILP_Total_s10)


  CC_founders_extinction_df <- CC_founders_extinction_df        %>%
  gather(key = Lever_Session, value = lever_presses, -DE_Subject, -DE_Strain, -DE_Sex) %>%
  mutate(Lever = gsub("^EX_([AI])LP_Total_s(\\d{2})$","\\1",Lever_Session),
         Session = as.numeric(gsub("^EX_([AI])LP_Total_s(\\d{2})$","\\2",Lever_Session))) %>%
  filter(!is.na(lever_presses), Session <= 8)
CC_founders_extinction_summary = CC_founders_extinction_df %>%
  group_by(DE_Strain, DE_Sex, Lever, Session) %>%
  summarize(mean = mean(lever_presses, na.rm = TRUE),
            n = length(which(!is.na(lever_presses))),
            sem = sd(lever_presses, na.rm = TRUE) / sqrt(length(which(!is.na(lever_presses))))) %>%
  mutate(ul = mean + sem, ll = mean - sem)

CC_founders_extinction_df
```

```{r}
CC_founders_extinction_slope = CC_founders_extinction_df %>%
  group_by(DE_Subject, Lever) %>%
  do(Sessions_lm = (lm(lever_presses ~ Session, data = .))) %>%
  broom::tidy(Sessions_lm) %>%
  filter(term == "Session") %>%
  select(DE_Subject, Lever, estimate) %>%
  mutate(Subject_Lever = paste(DE_Subject, Lever, sep = "_"))
CC_founders_extinction_slope = as.data.frame(CC_founders_extinction_slope)
row.names(CC_founders_extinction_slope) = CC_founders_extinction_slope$Subject_Lever

CC_founders_extinction_demodata = CC_founders_extinction_df%>%
  group_by(DE_Subject, Lever) %>%
  summarize(DE_Strain = unique(DE_Strain),
            DE_Sex = unique(DE_Sex)) %>%
  mutate(Subject_Lever = paste(DE_Subject, Lever, sep = "_"))
CC_founders_extinction_demodata = as.data.frame(CC_founders_extinction_demodata)
row.names(CC_founders_extinction_demodata) = CC_founders_extinction_demodata$Subject_Lever
CC_founders_extinction_demodata = CC_founders_extinction_demodata[,(-1 * which(colnames(CC_founders_extinction_demodata) == "Subject_Lever"))]
CC_founders_extinction_demodata$Slope = CC_founders_extinction_slope[row.names(CC_founders_extinction_demodata),"estimate"]
CC_founders_extinction_slope = CC_founders_extinction_demodata; rm(list = c("CC_founders_extinction_demodata"))
```

```{r}
CC_founders_extinction_slope_wide <- spread(CC_founders_extinction_slope, Lever, Slope)
CC_founders_extinction_slope_wide
```
ANOVA on slope of active lever presses during extinction session
```{r}
CC_founders_extinction_slope_wide %>%
  filter(!is.na(A)) %>%
  group_by(DE_Strain) %>%
  summarize(n = length(A))
anova(lm(A ~ DE_Strain, data = CC_founders_extinction_slope_wide))
```
Selecting the relevant columns for Reinstatement relevant trait calculations. Choosing the following variables:
'EX_ALP_Total_*last session': Active lever press for the final extinction session
'RI_Inf_Total_*s01_03'three extinction sessions.
Calculating %inc. in active lever press in reinstatement sessions relative to last extinction session.
Changing the format from wide to long format.

```{r}
CC_founders_RI_df = CC_founders_IVSA %>%
  select(DE_Subject,
         DE_Strain,
         DE_Sex,
         RI_Inf_Total_s01,
         RI_Inf_Total_s02,
         RI_Inf_Total_s03) %>%
  gather(key = Session, value = Infusions, -DE_Subject, -DE_Strain, -DE_Sex) %>%
  mutate(Session = as.numeric(gsub("^RI_Inf_Total_s(\\d{2})$","\\1",Session))) %>%
  filter(!is.na(Infusions))

CC_founders_RI_ALP_df = CC_founders_IVSA %>%
  select(DE_Subject,
         DE_Strain,
         DE_Sex,
         RI_ALP_Total_s01,
         RI_ALP_Total_s02,
         RI_ALP_Total_s03) %>%
  gather(key = Session, value = ALP, -DE_Subject, -DE_Strain, -DE_Sex) %>%
  mutate(Session = as.numeric(gsub("^RI_ALP_Total_s(\\d{2})$","\\1",Session))) %>%
  filter(!is.na(ALP))
CC_founders_RI_ALP_df = as.data.frame(CC_founders_RI_ALP_df)
CC_founders_RI_ALP_df
CC_founders_final_day_EX_ALP = CC_founders_extinction_df %>%
  group_by(DE_Subject) %>%
  filter(!is.na(lever_presses) & Lever == "A") %>%
  summarize(ALP = lever_presses[which(Session == max(Session))])
CC_founders_final_day_EX_ALP = as.data.frame(CC_founders_final_day_EX_ALP)
row.names(CC_founders_final_day_EX_ALP) = CC_founders_final_day_EX_ALP$DE_Subject 
CC_founders_final_day_EX_ALP 
CC_founders_RI_ALP_df
CC_founders_RI_ALP_df$final_EX_ALP = CC_founders_final_day_EX_ALP[CC_founders_RI_ALP_df$DE_Subject,"ALP"]
CC_founders_RI_ALP_df
CC_founders_RI_ALP_df = CC_founders_RI_ALP_df[which(CC_founders_RI_ALP_df$Session == 1),]
CC_founders_RI_ALP_df$percent_inc_final_EX = ((CC_founders_RI_ALP_df$ALP - CC_founders_RI_ALP_df$final_EX_ALP + 1) / (CC_founders_RI_ALP_df$final_EX_ALP + 1)) * 100
CC_founders_RI_ALP_df$ratio_final_EX = (CC_founders_RI_ALP_df$ALP + 1) / (CC_founders_RI_ALP_df$final_EX_ALP + 1)
CC_founders_RI_ALP_df
```

```{r}
anova(lm(percent_inc_final_EX ~ DE_Strain * DE_Sex, data = CC_founders_RI_ALP_df))

anova(lm(percent_inc_final_EX ~ DE_Strain, data = CC_founders_RI_ALP_df))
```


```{r}
total1 <- merge.data.frame(CC_founders_IVSA_filtered,CC_founders_extinction_slope_wide, by.x="DE_Subject", by.y="DE_Subject", all.x=TRUE)
total1
total <- merge.data.frame(total1,CC_founders_RI_ALP_df, by.x="DE_Subject", by.y="DE_Subject", all.x=TRUE)

total
names(total)[names(total) == 'A'] <- 'Slope_Extinction_Active_Lever'
names(total)[names(total) == 'percent_inc_final_EX'] <- 'RI_percent_final_EX'
total <- total [ , -which(names(total) %in% c("DE_Strain.y","DE_Sex.y","DE_Strain","DE_Sex"))]
names(total)[names(total) == 'DE_Strain.x'] <- 'DE_Strain'
names(total)[names(total) == 'DE_Sex.x'] <- 'DE_Sex'
CC_founders_IVSA_filtered <- total
```

Counting each cell to see if there's sufficient data to estimate heritability. 
This looks ready for downstream heritability analysis.

Saving these IVSA data for downstream use in the paper and in sharing on MPD.
```{r}
CC_founders_IVSA_filtered %>%
  filter(!is.na(Slope_Extinction_Active_Lever)) %>%
  group_by(DE_Strain) %>%
  summarize(n = length(Slope_Extinction_Active_Lever))
```

```{r}
CC_founders_IVSA_filtered %>%
  filter(!is.na(RI_percent_final_EX)) %>%
  group_by(DE_Strain) %>%
  summarize(n = length(RI_percent_final_EX))
```

```{r}
saveRDS(CC_founders_IVSA_filtered, paste("C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA/csna_project_4_behavior/data/IVSA_data_preprocessed_EX_RE_",current_date,".RDS",sep=""))
write.table(CC_founders_IVSA_filtered, paste("C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA/csna_project_4_behavior/data/IVSA_data_preprocessed_EX_RE_",current_date,".tsv",sep=""),
            sep = "\t", quote = FALSE, row.names = FALSE, col.names = TRUE)
```

# Reproducibility Information

## Software Used

### R session and OS information

```{r sessionInfo}
IVSA_data_prep_sessionInfo = sessionInfo()
IVSA_data_prep_sessionInfo
saveRDS(IVSA_data_prep_sessionInfo, paste("C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA/csna_project_4_behavior/output/IVSA_data_prep_sessionInfo_",current_date,".RDS", sep = ""), compress="xz")
```

## Document Control

This document was prepared using RMarkdown in RStudio.
