---
title: "IVSA Data: Preprocessing"
author: "Michael C. Saul (michael.saul [at] jax.org)"
adapted by: "Udita Datta (udita.datta [at] jax.org)"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: true
    fig_retina: true
    self_contained: true
    code_folding: show
    theme: flatly
    highlight: default
    df_print: paged
---

# Background

## Motivation

### Purpose of analysis

This analysis is performed to preprocess the intravenous self-administration data to:

1) Make sure all data are clean and deduplicated.
2) Make short and simple column names.
3) Clean up duplicate IDs in novelty data in a rational way.
4) Prepare data for cross-platform compatibility in both Python and R.

**These data were collected by RAs from the CSNA's behavioral core. Intellectual contributions will be assessed in discussions with Leona Gagnon.**

# Analysis

## Setup

### Working directory

All chunks will run in the working folder for this experiment.

```{r setup}
script_root_dir = "C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA"
knitr::opts_knit$set(root.dir = script_root_dir)
```

Getting current date

```{r}
current_date = format(Sys.time(),"%Y-%m-%d")
```

### R libraries

Calling R libraries necessary for this analysis.

```{r}
library("dplyr")
library("tidyverse")
library("lubridate")
library("readxl")
library("nlme")
library("ez")
```

## Importing data


```{r import_data, warning=FALSE}
# Taking the most recent datasets
data_path = paste(script_root_dir,"/preprocessed_IVSA_data/", sep = "")
IVSA_datasets = data.frame(files = list.files(path = data_path, pattern = "\\d{4}-\\d{2}-\\d{2}\\.RDS$"),
                           date = ymd(gsub("^.*_(\\d{4}-\\d{2}-\\d{2})\\.RDS$","\\1", 
                                           list.files(path = data_path, 
                                                      pattern = "\\d{4}-\\d{2}-\\d{2}\\.RDS$")), 
                                      tz = "America/New_York"),
                           stringsAsFactors = FALSE)
most_recent_datasets = paste(data_path, IVSA_datasets[which(IVSA_datasets$date == max(IVSA_datasets$date)),"files"], sep = "")

# IVSA
IVSA_raw_file = most_recent_datasets[grep("IVSA_raw", most_recent_datasets)]
IVSA_metadata_file = most_recent_datasets[grep("IVSA_metadata", most_recent_datasets)]
IVSA_raw = readRDS(IVSA_raw_file)
row.names(IVSA_raw) = IVSA_raw$DE_Subject
IVSA_metadata = readRDS(IVSA_metadata_file)
head(IVSA_raw)
```
Getting founders key file.

```{r founders_key_import, warning=FALSE}
founders_key = readRDS("C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA/csna_project_4_behavior/data/cc_do_founders_key.RDS")
row.names(founders_key) = founders_key$strain
founders_key[9,1] = "J:DO"
founders_key[9,2] = "DO"
founders_key[9,5] = "#000000"
founders_key[9,6] = "#000000"
row.names(founders_key) = founders_key$strain
head(IVSA_raw)
```

```{r filter_for_founders, warning=FALSE}
founders_IVSA = IVSA_raw %>%
  filter(DE_Strain %in% founders_key$strain)
#rm(list = c("IVSA_raw"))
founders_IVSA = as.data.frame(founders_IVSA)

founders_IVSA$DE_Strain = factor(founders_IVSA$DE_Strain,
                                 levels = founders_key$strain,
                                 ordered = TRUE)


Old_DO <- founders_IVSA %>% 
  filter(DE_birth_date_year == 2016,
         DE_Strain == "J:DO")


New_DO <- founders_IVSA %>% 
  filter(DE_birth_date_year != 2016,
         DE_Strain == "J:DO")

No_DO <- founders_IVSA %>% 
  filter(DE_Strain != "J:DO")

founders_IVSA <- merge(Old_DO,No_DO, all = TRUE)

ids <- c("A2063", "A2810", "A2418", "A3262", "A3261", "A3052", "A3257", "A3256", "A3051", "A2986")
rm.ids <- which(founders_IVSA$DE_Subject %in% ids)
founders_IVSA.p1  <- founders_IVSA[-rm.ids,]

founders_IVSA = founders_IVSA.p1

```
Getting CC key.

```{r}
CC_key = unique(IVSA_raw$DE_Strain)
CC_key = CC_key[grep("^CC\\d{3}/",CC_key)]
#CC_key = stringr::str_extract(CC_key, "^.{5}")
CC_key
CC_founders_IVSA_CC = IVSA_raw %>%
  filter(DE_Strain %in% c(CC_key))

CC_founders_IVSA_CC$DE_Strain = stringr::str_extract(CC_founders_IVSA_CC$DE_Strain, "^.{5}")
CC_founders_IVSA_CC
```

Merging old DO, founders and CC
```{r}
CC_founders_IVSA <- rbind(founders_IVSA, CC_founders_IVSA_CC)
```

The environment and datasets are ready for analysis.
# Analysis

## Acquisition

Excluding animals who were found dead who didn't acquire and changing the failed acquisition animals to 28 sessions.


```{r}
CC_founders_IVSA_filtered = CC_founders_IVSA %>%
  dplyr::filter(!(DE_Exit_Reason == "FD" & is.na(AQ_SessionsToAcquisition)))

CC_founders_IVSA_filtered[which(CC_founders_IVSA_filtered$DE_Exit_Reason == "IVSA Acquisition Only"),"AQ_SessionsToAcquisition"] = 28
CC_founders_IVSA_filtered
```
Looking at the rest of the data.

```{r}
CC_founders_IVSA_filtered[which(is.na(CC_founders_IVSA_filtered$AQ_SessionsToAcquisition)),]

subset(CC_founders_IVSA_filtered,is.na(CC_founders_IVSA_filtered$AQ_SessionsToAcquisition) & CC_founders_IVSA_filtered$DE_Exit_Reason == "Finished Pipeline")

```

Among founders: Animal `575` appears to have finished the pipeline, but the script could not calculate a sessions to acquisition criterion for it. All the other animals have exit reasons that should exclude them. Excluding these data.

A total of 1 DO and 5 CC individuals have finished the pipeline but the script could not calculate a sessions to acquisition criterion for it.

```{r}
CC_founders_IVSA_filtered = CC_founders_IVSA_filtered %>%
  filter(!is.na(AQ_SessionsToAcquisition))
```

Counting each cell to see if there's sufficient data to estimate heritability. Starting with `DR_Inf_Total_1p0mgkg`
```{r}
CC_founders_IVSA_filtered %>%
  filter(!is.na(DR_Inf_Total_1p0mgkg)) %>%
  group_by(DE_Strain) %>%
  summarize(n = length(DR_Inf_Total_1p0mgkg))
```

There appears to be adequate sampling for this dataset. Counting each cell of the `AQ_SessionsToAcquisition` data.

```{r}
CC_founders_IVSA_filtered %>%
  filter(!is.na(AQ_SessionsToAcquisition)) %>%
  group_by(DE_Strain) %>%
  summarize(n = length(AQ_SessionsToAcquisition))
```
Counting each cell of the `DR_Inf_Total_AUC` data.

```{r}
CC_founders_IVSA_filtered %>%
  filter(!is.na(DR_Inf_Total_AUC)) %>%
  group_by(DE_Strain) %>%
  summarize(n = length(DR_Inf_Total_AUC))
```


This looks ready for downstream heritability analysis.

Saving these IVSA data for downstream use in the paper and in sharing on MPD.


```{r}
saveRDS(CC_founders_IVSA_filtered, paste("C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA/csna_project_4_behavior/data/IVSA_data_preprocessed_AQ_DR_",current_date,".RDS",sep=""))
write.table(CC_founders_IVSA_filtered, paste("C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA/csna_project_4_behavior/data/IVSA_data_preprocessed_AQ_DR_",current_date,".tsv",sep=""),
            sep = "\t", quote = FALSE, row.names = FALSE, col.names = TRUE)
```

# Reproducibility Information

## Software Used

### R session and OS information

```{r sessionInfo}
IVSA_data_prep_sessionInfo = sessionInfo()
IVSA_data_prep_sessionInfo
saveRDS(IVSA_data_prep_sessionInfo, paste("C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA/csna_project_4_behavior/output/IVSA_data_prep_sessionInfo_",current_date,".RDS", sep = ""), compress="xz")
```

## Document Control

This document was prepared using RMarkdown in RStudio.
