---
title: "Founders and CC Intravenous Self-Administration Data: Heritability and Plotting"
author: "Michael C. Saul (michael.saul [at] jax.org)"
modified by: "Udita Datta (udita.datta [at] jax.org)"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: true
    fig_retina: true
    self_contained: true
    code_folding: show
    theme: flatly
    highlight: default
    df_print: paged
---

# Background

## Motivation

### Purpose of analysis

This analysis is performed to process the intravenous self-administration data and calculate:

1) Heritability
2) Strain-by-sex means and standard errors for plotting

**These data were collected by the CSNA research assistants. Ultimate authorship of the resultant publications will be determined in discussion with Leona Gagnon.** The preprocessing script `IVSA_import_preprocess.Rmd` was used to filter the data to only those data requiring reporting.

### Note about this analysis

The data have a floor effect (animals that took no cocaine) for infusions and a ceiling effect (animals that did not acquire self-administration) for sessions to acquisition. The appropriateness of a conventional linear model should be discussed with Vivek Philip. It's possible that we should fit a Tobit regression.

# Analysis

## Setup

### Working directory

All chunks will run in the working folder for this experiment.

```{r setup}
script_root_dir = "C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA"
knitr::opts_knit$set(root.dir = script_root_dir)
```

Getting current date

```{r}
current_date = format(Sys.time(),"%Y-%m-%d")
```

### R libraries

Calling R libraries necessary for this analysis.

```{r}
library("tidyverse")
library("lubridate")
library("cowplot")
library("MASS")
library("ggbeeswarm")
```

## Importing data

## Importing and displaying raw intravenous self-administration data

Getting data from the filtering script and taking out the `DE_` part of the header

```{r}
IVSA_data = readRDS(paste("C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA/csna_project_4_behavior/data/IVSA_data_preprocessed_AQ_DR_2020-12-17.RDS", sep = ""))
colnames(IVSA_data) = gsub("^DE_","",colnames(IVSA_data))
row.names(IVSA_data) = as.character(IVSA_data$Subject)
IVSA_data
```

Displaying raw data.

```{r}
IVSA_data$failed_aq = ifelse(IVSA_data$Exit_Reason == "IVSA Acquisition Only", 2, 1)
View(IVSA_data)
```

Getting founders strain key. Getting the founders only (no CC or DO) for heritabiliy calculations. 
```{r founders_key_import, warning=FALSE}
founders_key = readRDS("C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA/csna_project_4_behavior/data/cc_do_founders_key.RDS")
row.names(founders_key) = founders_key$strain
IVSA_data = IVSA_data %>%
  filter(Strain %in% founders_key$strain)
#rm(list = c("IVSA_raw"))
IVSA_data = as.data.frame(IVSA_data)
```


### Fitting ANOVA model for intravenous self-administration dose-response 1.0 mg/kg data

Fitting the model with genetic background (strain), sex, and strain x sex interactions.

```{r}
anova(lm(AQ_SessionsToAcquisition ~ Strain * Sex, data = IVSA_data))
anova(lm(DR_Inf_Total_1p0mgkg ~ Strain * Sex, data = IVSA_data))
anova(lm(DR_Inf_Total_AUC ~ Strain * Sex, data = IVSA_data))
```


Strain is the only significant factor here. Looking at the residuals of the simplest model with just this in it first.

```{r}
anova(lm(AQ_SessionsToAcquisition ~ Strain, data = IVSA_data))
anova(lm(DR_Inf_Total_1p0mgkg ~ Strain, data = IVSA_data))
anova(lm(DR_Inf_Total_AUC ~ Strain, data = IVSA_data))
```
Plotting residuals with only strain in the model.
```{r}
IVSA_resid = IVSA_data[!sapply(IVSA_data, function(x) any(is.na(x)))]
IVSA_resid
IVSA_resid$DR_Inf_Total_1p0mgkg_Resid = resid(lm(DR_Inf_Total_1p0mgkg ~ Strain, data = IVSA_resid))
IVSA_resid$Strain = factor(IVSA_resid$Strain,
                                           levels = founders_key$strain,
                                           ordered = TRUE)

ggplot(data = IVSA_resid, aes(x = Strain, y = DR_Inf_Total_1p0mgkg_Resid, color = Strain, shape = Sex)) +
  geom_hline(yintercept = 0, col = "#999999") +
  geom_beeswarm(size = 2) + 
  scale_color_manual(values = founders_key$collaborative_cross_color_broman) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(16,23))
```

```{r}
qqnorm(IVSA_data$DR_Inf_Total_1p0mgkg, pch = 1, frame = FALSE)
qqline(IVSA_data$DR_Inf_Total_1p0mgkg, col = "steelblue", lwd = 2)
```

The residuals indicate heteroscedasticity. 129 doesn't respond, so not worrying about those points. *NOTE: Heritability should be calculated both with 129 and without to prevent upward bias.* However, the PWK residuals appear to be at least 2.5-fold greater in magnitude than the WSB residuals. Doing a Box-Cox to see if a response variable transformation is merited.

```{r}
boxcox(lm((DR_Inf_Total_1p0mgkg + 1) ~ Strain, data = IVSA_resid))
```

The Box-Cox log-likelihood is highest at about 1/3, which implies that a cube root transform will be advantageous.

```{r}
boxcox(lm(1 + (DR_Inf_Total_1p0mgkg ^ (1/3)) ~ Strain, data = IVSA_resid))
```

The Box-Cox log-likelihood is centered closer to 1. Plotting residuals.

```{r}
IVSA_resid$DR_Inf_Total_1p0mgkg_Resid_cuberoot = resid(lm((DR_Inf_Total_1p0mgkg ^ (1/3)) ~ Strain, data = IVSA_resid))

ggplot(data = IVSA_resid, aes(x = Strain, y = DR_Inf_Total_1p0mgkg_Resid_cuberoot, color = Strain, shape = Sex)) +
  geom_hline(yintercept = 0, col = "#999999") +
  geom_beeswarm(size = 2) + 
  scale_color_manual(values = founders_key$collaborative_cross_color_broman) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(16,23))
```

These residuals are better. Doing model selection based upon this transform.

```{r}
AICs = c()
AICs["AIC_inter"] = AIC(lm((DR_Inf_Total_1p0mgkg ^ (1/3))  ~ Strain + Sex + Strain:Sex, data = IVSA_data))
AICs["AIC_nointer"] = AIC(lm((DR_Inf_Total_1p0mgkg ^ (1/3))  ~ Strain + Sex, data = IVSA_data))
AICs["AIC_nosex"] = AIC(lm((DR_Inf_Total_1p0mgkg ^ (1/3))  ~ Strain, data = IVSA_data))
AICs["AIC_nostrain"] = AIC(lm((DR_Inf_Total_1p0mgkg ^ (1/3))  ~ 1, data = IVSA_data))

AICs

BICs = c()
BICs["BIC_inter"] = BIC(lm((DR_Inf_Total_1p0mgkg ^ (1/3))  ~ Strain + Sex + Strain:Sex, data = IVSA_data))
BICs["BIC_nointer"] = BIC(lm((DR_Inf_Total_1p0mgkg ^ (1/3))  ~ Strain + Sex, data = IVSA_data))
BICs["BIC_nosex"] = BIC(lm((DR_Inf_Total_1p0mgkg ^ (1/3))  ~ Strain, data = IVSA_data))
BICs["BIC_nostrain"] = BIC(lm((DR_Inf_Total_1p0mgkg ^ (1/3))  ~ 1, data = IVSA_data))

BICs
```

The best model for heritability according to both AIC and BIC contain only strain with no sex and strain:sex interaction factors. The residuals for this model have been plotted above. Showing the ANOVA table for this model.

```{r}
anova(lm((DR_Inf_Total_1p0mgkg ^ (1/3))  ~ Strain, data = IVSA_data))
anova(lm((DR_Inf_Total_1p0mgkg ^ (1/3))  ~ Strain, data = IVSA_data))$`Pr(>F)`[1]
```

This ANOVA model will be used for downstream heritability calculations.

### Calculating heritability for dose-response 1.0 mg/kg data

Calculating heritability using the ANOVA table from the best fitting model. Using the following equation:

$h^2=\displaystyle{{MS_{strain}}\over{MS_{strain} + (n_{mean} - 1) * MS_{resid}}}$

where $MS_{strain}$ is the mean square of the strain differences, $MS_{resid}$ is the mean square of the residuals, and $n_{mean}$ is the mean sample size of a strain. Starting with the IVSA data that include 129.

```{r}
mean_n = nrow(IVSA_data) / length(unique(IVSA_data$Strain))
anova_DR_1p0mgkg = anova(lm((DR_Inf_Total_1p0mgkg ^ (1/3))  ~ Strain, data = IVSA_data))
anova_DR_1p0mgkg_df = as.data.frame(anova_DR_1p0mgkg)
h2_DR_1p0mgkg_df = anova_DR_1p0mgkg_df["Strain","Mean Sq"] / (anova_DR_1p0mgkg_df["Strain","Mean Sq"] + ((mean_n - 1) * anova_DR_1p0mgkg_df["Residuals","Mean Sq"]))
h2_DR_1p0mgkg_df
```

This heritability of `r round(h2_DR_1p0mgkg_df,2)` is higher than previous estimates. It should be noted that this is reasonably close to the estimate of cocaine use disorder heritability that's observed in humans.

Calculating the heritability without 129.

```{r}
IVSA_data_no_129 = IVSA_data[which(IVSA_data$Strain != "129S1/SvImJ"),]
IVSA_data_no_129$Strain = as.character(IVSA_data_no_129$Strain)
mean_n = nrow(IVSA_data_no_129) / length(unique(IVSA_data_no_129$Strain))
anova_DR_1p0mgkg_no_129 = anova(lm((DR_Inf_Total_1p0mgkg ^ (1/3))  ~ Strain, data = IVSA_data_no_129))
anova_DR_1p0mgkg_no_129_df = as.data.frame(anova_DR_1p0mgkg_no_129)
h2_DR_1p0mgkg_df_no_129 = anova_DR_1p0mgkg_no_129_df["Strain","Mean Sq"] / (anova_DR_1p0mgkg_no_129_df["Strain","Mean Sq"] + ((mean_n - 1) * anova_DR_1p0mgkg_no_129_df["Residuals","Mean Sq"]))
h2_DR_1p0mgkg_df_no_129
```

129 appreciably drives the heritability; the number of `r round(h2_DR_1p0mgkg_df_no_129,2)` is a significant drop. 

### Plotting strain x sex means for total infusions at FR1 1.0 mg/kg

Plotting the strain:sex information for these data. The calculations will be back-transformed.

```{r}
IVSA_DR_1p0mgkg_summary = IVSA_data %>%
  group_by(Strain, Sex) %>%
  summarize(Strain_Sex_mean = mean((DR_Inf_Total_1p0mgkg ^ (1/3)), na.rm = TRUE),
            Strain_Sex_sem = sd((DR_Inf_Total_1p0mgkg ^ (1/3)), na.rm = TRUE) / length(which(!is.na((DR_Inf_Total_1p0mgkg ^ (1/3))))),
            Strain_Sex_n = length(which(!is.na(log10(AQ_SessionsToAcquisition))))) %>%
  mutate(Strain_Sex = paste(Strain, gsub("^([MF]).*$", "\\1", Sex), sep = " "),
         Strain_Sex_backmean = Strain_Sex_mean ^ 3,
         Strain_Sex_ul = (Strain_Sex_mean + Strain_Sex_sem) ^ 3, 
         Strain_Sex_ll = (Strain_Sex_mean - Strain_Sex_sem) ^ 3)
IVSA_DR_1p0mgkg_summary = as.data.frame(IVSA_DR_1p0mgkg_summary)
IVSA_DR_1p0mgkg_summary$Strain_Sex = factor(IVSA_DR_1p0mgkg_summary$Strain_Sex,
                                                levels = paste(rep(founders_key$strain, each = 2), rep(c(" F", " M"), times = 8), sep = ""),
                                                ordered = TRUE)
IVSA_DR_1p0mgkg_summary$Strain = factor(IVSA_DR_1p0mgkg_summary$Strain,
                                            levels = founders_key$strain,
                                            ordered = TRUE)

founders_IVSA_DR_1p0mgkg_plot = ggplot(data = IVSA_DR_1p0mgkg_summary, aes(x = Strain_Sex, y = Strain_Sex_backmean, color = Strain, group = Strain)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = Strain_Sex_ll, ymax = Strain_Sex_ul), width = 0.5, size = 1) +
  scale_color_manual(values = founders_key$collaborative_cross_color_broman) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position="none") +
  xlab(NULL) +
  ylab("FR1 Total Infusions at 1.0 mg/kg") +
  ggtitle("IVSA: 1.0 mg/kg (back-transformed from cube root)")

founders_IVSA_DR_1p0mgkg_plot
```

Saving plot

```{r}
saveRDS(founders_IVSA_DR_1p0mgkg_plot, "C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA/csna_project_4_behavior/output/founders_IVSA_DR_1p0mgkg_plot_sex_plot.RDS")
```

### Fitting ANOVA model for intravenous sessions to acquisition for self-administraion at 1.0 mg/kg

Fitting the model with genetic background (strain), sex, and strain x sex interactions.

```{r}
anova(lm(AQ_SessionsToAcquisition ~ Strain * Sex, data = IVSA_data))
```

Again, strain is the only significant factor in this dataset. Plotting residuals of model with only this factor.

Strain is the only significant factor here. Looking at the residuals of the simplest model with just this in it first.

```{r}
IVSA_resid$AQ_SessionsToAcquisition_Resid = resid(lm(AQ_SessionsToAcquisition ~ Strain, data = IVSA_resid))

ggplot(data = IVSA_resid, aes(x = Strain, y = AQ_SessionsToAcquisition_Resid, color = Strain, shape = Sex)) +
  geom_hline(yintercept = 0, col = "#999999") +
  geom_beeswarm(size = 2) + 
  scale_color_manual(values = founders_key$collaborative_cross_color_broman) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(16,23))
```

This fit doesn't look good. Trying Box-Cox to see if a response variable transform may be merited.

```{r}
boxcox(lm(AQ_SessionsToAcquisition ~ Strain, data = IVSA_resid))
```

The log-likelihood centered at 0 implies that a log-transform may help with the non-normal residuals.

```{r}
IVSA_resid$AQ_SessionsToAcquisition_Resid_log10 = resid(lm(log10(AQ_SessionsToAcquisition) ~ Strain, data = IVSA_resid))

ggplot(data = IVSA_resid, aes(x = Strain, y = AQ_SessionsToAcquisition_Resid_log10, color = Strain, shape = Sex)) +
  geom_hline(yintercept = 0, col = "#999999") +
  geom_beeswarm(size = 2) + 
  scale_color_manual(values = founders_key$collaborative_cross_color_broman) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(16,23))
```

This looks a little better. Trying Box-Cox to see if the log-likelihood is centered at 1 with the transformed data.

```{r}
boxcox(lm(log10(AQ_SessionsToAcquisition) ~ Strain, data = IVSA_resid))
```

This looks acceptable, though not ideal. Proceeding with model selection using the log10 transformation.

```{r}
AICs = c()
AICs["AIC_inter"] = AIC(lm(log10(AQ_SessionsToAcquisition) ~ Strain + Sex + Strain:Sex, data = IVSA_data))
AICs["AIC_nointer"] = AIC(lm(log10(AQ_SessionsToAcquisition)  ~ Strain + Sex, data = IVSA_data))
AICs["AIC_nosex"] = AIC(lm(log10(AQ_SessionsToAcquisition)  ~ Strain, data = IVSA_data))
AICs["AIC_nostrain"] = AIC(lm(log10(AQ_SessionsToAcquisition) ~ 1, data = IVSA_data))

AICs

BICs = c()
BICs["BIC_inter"] = BIC(lm(log10(AQ_SessionsToAcquisition)  ~ Strain + Sex + Strain:Sex, data = IVSA_data))
BICs["BIC_nointer"] = BIC(lm(log10(AQ_SessionsToAcquisition)  ~ Strain + Sex, data = IVSA_data))
BICs["BIC_nosex"] = BIC(lm(log10(AQ_SessionsToAcquisition) ~ Strain, data = IVSA_data))
BICs["BIC_nostrain"] = BIC(lm(log10(AQ_SessionsToAcquisition) ~ 1, data = IVSA_data))

BICs
```

The one-way ANOVA with just the strain factor appears to be the best model. Displaying the ANOVA table.

```{r}
anova(lm(log10(AQ_SessionsToAcquisition) ~ Strain, data = IVSA_data))
anova(lm(log10(AQ_SessionsToAcquisition) ~ Strain, data = IVSA_data))$`Pr(>F)`[1]
```

### Calculating heritability for sessions to acquisition

Calculating heritability using the ANOVA table from the best fitting model. Using the following equation:

$h^2=\displaystyle{{MS_{strain}}\over{MS_{strain} + (n_{mean} - 1) * MS_{resid}}}$

where $MS_{strain}$ is the mean square of the strain differences, $MS_{resid}$ is the mean square of the residuals, and $n_{mean}$ is the mean sample size of a strain.

Again, 129 has a ceiling effect, which would upwardly bias heritability calculations. Doing the calculation with and without 129.

```{r}
mean_n = nrow(IVSA_data) / length(unique(IVSA_data$Strain))
anova_AQ_SessionsToAcquisition = anova(lm(log10(AQ_SessionsToAcquisition) ~ Strain, data = IVSA_data))
IVSA_data
anova_AQ_SessionsToAcquisition_df = as.data.frame(anova_AQ_SessionsToAcquisition)
h2_AQ_SessionsToAcquisition = anova_AQ_SessionsToAcquisition_df["Strain","Mean Sq"] / (anova_AQ_SessionsToAcquisition_df["Strain","Mean Sq"] + ((mean_n - 1) * anova_AQ_SessionsToAcquisition_df["Residuals","Mean Sq"]))
h2_AQ_SessionsToAcquisition
```

This heritability of `r round(h2_AQ_SessionsToAcquisition,2)` is also higher than previous estimates.

Doing the calculation without 129.

```{r}
mean_n = nrow(IVSA_data_no_129) / length(unique(IVSA_data_no_129$Strain))
anova_AQ_SessionsToAcquisition_no_129 = anova(lm(log10(AQ_SessionsToAcquisition) ~ Strain, data = IVSA_data_no_129))
anova_AQ_SessionsToAcquisition_no_129_df = as.data.frame(anova_AQ_SessionsToAcquisition_no_129)
h2_AQ_SessionsToAcquisition_no_129 = anova_AQ_SessionsToAcquisition_no_129_df["Strain","Mean Sq"] / (anova_AQ_SessionsToAcquisition_no_129_df["Strain","Mean Sq"] + ((mean_n - 1) * anova_AQ_SessionsToAcquisition_no_129_df["Residuals","Mean Sq"]))
h2_AQ_SessionsToAcquisition_no_129
```

This heritability of `r round(h2_AQ_SessionsToAcquisition_no_129,2)` is a good magnitude.

### Plotting strain x sex means for sessions to acquisition

Plotting the strain:sex information for these data. The calculations will be back-transformed.

```{r}
IVSA_AQ_SessionsToAcquisition_summary = IVSA_data %>%
  group_by(Strain, Sex) %>%
  summarize(Strain_Sex_mean = mean(log10(AQ_SessionsToAcquisition), na.rm = TRUE),
            Strain_Sex_sem = sd(log10(AQ_SessionsToAcquisition), na.rm = TRUE) / length(which(!is.na(log10(AQ_SessionsToAcquisition)))),
            Strain_Sex_n = length(which(!is.na(log10(AQ_SessionsToAcquisition))))) %>%
  mutate(Strain_Sex = paste(Strain, gsub("^([MF]).*$", "\\1", Sex), sep = " "),
         Strain_Sex_backmean = 10 ^ Strain_Sex_mean,
         Strain_Sex_ul = 10 ^ (Strain_Sex_mean + Strain_Sex_sem), 
         Strain_Sex_ll = 10 ^ (Strain_Sex_mean - Strain_Sex_sem))
IVSA_AQ_SessionsToAcquisition_summary = as.data.frame(IVSA_AQ_SessionsToAcquisition_summary)
IVSA_AQ_SessionsToAcquisition_summary$Strain_Sex = factor(IVSA_AQ_SessionsToAcquisition_summary$Strain_Sex,
                                                levels = paste(rep(founders_key$strain, each = 2), rep(c(" F", " M"), times = 8), sep = ""),
                                                ordered = TRUE)
IVSA_AQ_SessionsToAcquisition_summary$Strain = factor(IVSA_AQ_SessionsToAcquisition_summary$Strain,
                                            levels = founders_key$strain,
                                            ordered = TRUE)

IVSA_AQ_SessionsToAcquisition_plot = ggplot(data = IVSA_AQ_SessionsToAcquisition_summary, aes(x = Strain_Sex, y = Strain_Sex_backmean, color = Strain, group = Strain)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = Strain_Sex_ll, ymax = Strain_Sex_ul), width = 0.5, size = 1) +
  scale_color_manual(values = founders_key$collaborative_cross_color_broman) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position="none") +
  xlab(NULL) +
  ylab("Sessions to Acquisition") +
  ggtitle("IVSA: Sessions to Acquisition (FR1 1.0 mg/kg)")

IVSA_AQ_SessionsToAcquisition_plot
```

Plotting individual points for IVSA infusions at FR-1 1.0 mg/kg

```{r}
IVSA_data$Strain_Sex = factor(paste(as.character(IVSA_data$Strain),
                                    as.character(IVSA_data$Sex),
                                    sep = " "),
                              levels = paste(rep(founders_key$strain, each = 2), rep(c("F","M"), times = 8), sep = " "),
                              ordered = TRUE)
IVSA_data$Strain = factor(IVSA_data$Strain, levels = founders_key$strain, ordered = TRUE)

founders_FR1_individual_points_plot = ggplot(data = IVSA_data, aes(x = Strain_Sex, y = DR_Inf_Total_1p0mgkg, color = Strain, shape = Sex)) +
  geom_beeswarm(size = 2) +
  scale_color_manual(values = c(founders_key$collaborative_cross_color_broman)) +
  scale_shape_manual(values = c(19,17)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position="none") +
  xlab(NULL) +
  ylab("FR1 Total Infusions at 1.0 mg/kg") +
  ggtitle("IVSA: 1.0 mg/kg Infusions")
founders_FR1_individual_points_plot
```

Plotting individual points for IVSA sessions to acquisition

```{r}
founders_StA_individual_points_plot = ggplot(data = IVSA_data, aes(x = Strain_Sex, y = AQ_SessionsToAcquisition, color = Strain, shape = Sex)) +
  geom_beeswarm(size = 2) +
  scale_color_manual(values = c(founders_key$collaborative_cross_color_broman)) +
  scale_shape_manual(values = c(19,17)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position="none") +
  xlab(NULL) +
  ylab("Sessions to Acquisition") +
  ggtitle("IVSA: Sessions to Acquisition (FR1 1.0 mg/kg)")
founders_StA_individual_points_plot
```

Saving plot

```{r}
saveRDS(IVSA_AQ_SessionsToAcquisition_plot, "C:/Users/dattau/Documents/R_IVSA_Troy/IVSA/IVSA/csna_project_4_behavior/output/IVSA_AQ_SessionsToAcquisition_strain_sex_plot.RDS")
```

### Fitting ANOVA model for DR_Inf_Total_AUC (Area under the curve for Dose Response phase of self-administraion

Fitting the model with genetic background (strain), sex, and strain x sex interactions.

```{r}
anova(lm(DR_Inf_Total_AUC ~ Strain * Sex, data = IVSA_data))
```

Again, strain and strain by sex interaction are significant  in this dataset. Plotting residuals of model with these factors.



```{r}
founders_key_no_129 <- founders_key[which(founders_key$strain != "129S1/SvImJ"),]
IVSA_resid = IVSA_data %>% 
  filter(!is.na(DR_Inf_Total_AUC))
IVSA_resid$DR_Inf_Total_AUC_Resid = resid(lm(DR_Inf_Total_AUC ~ Strain, data = IVSA_resid))

ggplot(data = IVSA_resid, aes(x = Strain, y = DR_Inf_Total_AUC, color = Strain, shape = Sex)) +
  geom_hline(yintercept = 500, col = "#999999") +
  geom_beeswarm(size = 2) + 
  scale_color_manual(values = founders_key_no_129$collaborative_cross_color_broman) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(16,23))
```

This fit doesn't look good. Trying Box-Cox to see if a response variable transform may be merited.

```{r}
boxcox(lm(DR_Inf_Total_AUC ~ Strain, data = IVSA_resid))
```

The lambda centered close 0.5 implies that a sqrt transform may help with the non-normal residuals.

```{r}
founders_key_no_129 <- founders_key[which(founders_key$strain != "129S1/SvImJ"),]
IVSA_resid$DR_Inf_Total_AUC_sqrt = resid(lm(sqrt(DR_Inf_Total_AUC) ~ Strain, data = IVSA_resid))

ggplot(data = IVSA_resid, aes(x = Strain , y = DR_Inf_Total_AUC_sqrt, color = Strain, shape = Sex)) +
  geom_hline(yintercept = 0, col = "#999999") +
  geom_beeswarm(size = 2) + 
  scale_color_manual(values = founders_key_no_129$collaborative_cross_color_broman) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(16,23))
```

This looks a little better. Trying Box-Cox to see if the log-likelihood is centered at 1 with the transformed data.

```{r}
boxcox(lm(sqrt(DR_Inf_Total_AUC) ~ Strain * Sex, data = IVSA_resid))
```

This looks acceptable. Proceeding with model selection using the sqrt transformation.

```{r}
AICs = c()
AICs["AIC_inter"] = AIC(lm(sqrt(DR_Inf_Total_AUC) ~ Strain + Sex + Strain:Sex, data = IVSA_data))
AICs["AIC_nointer"] = AIC(lm(sqrt(DR_Inf_Total_AUC)  ~ Strain + Sex, data = IVSA_data))
AICs["AIC_nosex"] = AIC(lm(sqrt(DR_Inf_Total_AUC)  ~ Strain, data = IVSA_data))
AICs["AIC_nostrain"] = AIC(lm(sqrt(DR_Inf_Total_AUC) ~ 1, data = IVSA_data))

AICs

BICs = c()
BICs["BIC_inter"] = BIC(lm(sqrt(DR_Inf_Total_AUC)  ~ Strain + Sex + Strain:Sex, data = IVSA_data))
BICs["BIC_nointer"] = BIC(lm(sqrt(DR_Inf_Total_AUC)  ~ Strain + Sex, data = IVSA_data))
BICs["BIC_nosex"] = BIC(lm(sqrt(DR_Inf_Total_AUC) ~ Strain, data = IVSA_data))
BICs["BIC_nostrain"] = BIC(lm(sqrt(DR_Inf_Total_AUC) ~ 1, data = IVSA_data))

BICs
```

The one-way ANOVA with just the strain factor appears to be the best model. Displaying the ANOVA table.

```{r}
anova(lm(sqrt(DR_Inf_Total_AUC) ~ Strain * Sex, data = IVSA_data))
anova(lm(sqrt(DR_Inf_Total_AUC) ~ Strain * Sex, data = IVSA_data))$`Pr(>F)`[1]
```

### Calculating heritability for dose response AUC

Calculating heritability using the ANOVA table from the best fitting model. Using the following equation:

$h^2=\displaystyle{{MS_{strain}}\over{MS_{strain} + (n_{mean} - 1) * MS_{resid}}}$

where $MS_{strain}$ is the mean square of the strain differences, $MS_{resid}$ is the mean square of the residuals, and $n_{mean}$ is the mean sample size of a strain.


```{r}
mean_n = nrow(IVSA_data) / length(unique(IVSA_data$Strain))
anova_DR_Inf_Total_AUC = anova(lm(sqrt(DR_Inf_Total_AUC) ~ Strain, data = IVSA_data))
anova_DR_Inf_Total_AUC_df = as.data.frame(anova_DR_Inf_Total_AUC)
h2_DR_Inf_Total_AUC = anova_DR_Inf_Total_AUC_df["Strain","Mean Sq"] / (anova_DR_Inf_Total_AUC_df["Strain","Mean Sq"] + ((mean_n - 1) * anova_DR_Inf_Total_AUC_df["Residuals","Mean Sq"]))
h2_DR_Inf_Total_AUC
```


```{r}
mean_n = nrow(IVSA_data) / (length(unique(IVSA_data$Strain)) * length(unique(IVSA_data$Sex)))
anova_DR_Inf_Total_AUC = anova(lm(sqrt(DR_Inf_Total_AUC) ~ Strain * Sex, data = IVSA_data))
anova_DR_Inf_Total_AUC_df = as.data.frame(anova_DR_Inf_Total_AUC)
h2_DR_Inf_Total_AUC = (anova_DR_Inf_Total_AUC_df["Strain","Mean Sq"]-anova_DR_Inf_Total_AUC_df["Strain:Sex","Mean Sq"]) /(2 * (anova_DR_Inf_Total_AUC_df["Strain","Mean Sq"] + (mean_n -1) * anova_DR_Inf_Total_AUC_df["Residuals","Mean Sq"]))
h2_DR_Inf_Total_AUC

#h2 = round((msb-msi)/(s*(msb+(n-1)*mse)),2)
save(list=ls(), file="IVSA.Rdata")
```

This heritability of `r round(h2_DR_Inf_Total_AUC,2)` is also higher than previous estimates.

Doing the calculation without 129.

```{r}
mean_n = nrow(IVSA_data_no_129) / length(unique(IVSA_data_no_129$Strain))
anova_DR_Inf_Total_AUC_no_129 = anova(lm(sqrt(DR_Inf_Total_AUC) ~ Strain, data = IVSA_data_no_129))
anova_DR_Inf_Total_AUC_no_129_df = as.data.frame(anova_DR_Inf_Total_AUC_no_129)
h2_DR_Inf_Total_AUC_no_129 = anova_DR_Inf_Total_AUC_no_129_df["Strain","Mean Sq"] / (anova_DR_Inf_Total_AUC_no_129_df["Strain","Mean Sq"] + ((mean_n - 1) * anova_DR_Inf_Total_AUC_no_129_df["Residuals","Mean Sq"]))
h2_DR_Inf_Total_AUC_no_129
```



### Plotting strain x sex means for sessions to acquisition

Plotting the strain:sex information for these data. The calculations will be back-transformed.

```{r}
founders_key_no_129 <- founders_key[which(founders_key$strain != "129S1/SvImJ"),]
IVSA_DR_Inf_Total_AUC_summary = IVSA_data %>% filter(Strain != "129S1/SvImJ")%>%
  group_by(Strain, Sex) %>%
  summarize(Strain_Sex_mean = mean(sqrt(DR_Inf_Total_AUC), na.rm = TRUE),
            Strain_Sex_sem = sd(sqrt(DR_Inf_Total_AUC), na.rm = TRUE) / length(which(!is.na(sqrt(DR_Inf_Total_AUC)))),
            Strain_Sex_n = length(which(!is.na(sqrt(DR_Inf_Total_AUC))))) %>%
  mutate(Strain_Sex = paste(Strain, gsub("^([MF]).*$", "\\1", Sex), sep = " "),
         Strain_Sex_backmean =  Strain_Sex_mean ^ 2 ,
         Strain_Sex_ul = (Strain_Sex_mean + Strain_Sex_sem)^ 2, 
         Strain_Sex_ll = (Strain_Sex_mean - Strain_Sex_sem)^ 2)
IVSA_DR_Inf_Total_AUC_summary = as.data.frame(IVSA_DR_Inf_Total_AUC_summary)
IVSA_DR_Inf_Total_AUC_summary$Strain_Sex = factor(IVSA_DR_Inf_Total_AUC_summary$Strain_Sex,
                                                levels = paste(rep(founders_key_no_129$strain, each = 2), rep(c(" F", " M"), times = 6), sep = ""),
                                                ordered = TRUE)
IVSA_DR_Inf_Total_AUC_summary$Strain = factor(IVSA_DR_Inf_Total_AUC_summary$Strain,
                                            levels = founders_key_no_129$strain,
                                            ordered = TRUE)

IVSA_DR_Inf_Total_AUC_plot = ggplot(data = IVSA_DR_Inf_Total_AUC_summary, aes(x = Strain_Sex, y = Strain_Sex_backmean, color = Strain, group = Strain)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = Strain_Sex_ll, ymax = Strain_Sex_ul), width = 0.5, size = 1) +
  scale_color_manual(values = founders_key_no_129$collaborative_cross_color_broman) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position="none") +
  xlab(NULL) +
  ylab("Dose-Response Infusions AUC") +
  ggtitle("IVSA: Dose-Response Infusions AUC")

IVSA_DR_Inf_Total_AUC_plot
```

Plotting individual points for IVSA infusions at FR-1 1.0 mg/kg


Saving plot

```{r}
saveRDS(IVSA_AQ_SessionsToAcquisition_plot, "./output/IVSA_AQ_SessionsToAcquisition_strain_sex_plot.RDS")
```

## Document Control

This document was prepared using RMarkdown in RStudio.
