---
title: "Fig4_IVSA_Manuscript"
author: "Udita Datta"
date: "12/10/2020"
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: true
    fig_retina: true
    self_contained: true
    code_folding: show
    theme: flatly
    highlight: default
    df_print: paged
---


Seting up Working directory


```{r setup, warning=FALSE}
script_root_dir = "C:/Users/dattau/Documents/IVSA_Proj_org/IVSA/IVSA"
knitr::opts_knit$set(root.dir = script_root_dir)
```

### R libraries

Calling R libraries necessary for this analysis.

```{r libraries, warning=FALSE}
library("MASS")
library("cowplot")
library("lubridate")
library("ggbeeswarm")
library("readxl")
library("tidyr")
library("dplyr")
library("DT")
library("ANOM")
library("multcomp")
library("ez")
```

## Dataset Setup

### Importing data

The IVSA and novelty data have already been preprocessed using the `IVSA_novelty_import_preprocess.Rmd` script.

```{r import_data, warning=FALSE}
# Taking the most recent datasets
data_path = paste(script_root_dir,"/preprocessed_IVSA_data/", sep = "")
IVSA_datasets = data.frame(files = list.files(path = data_path, pattern = "\\d{4}-\\d{2}-\\d{2}\\.RDS$"),
                           date = ymd(gsub("^.*_(\\d{4}-\\d{2}-\\d{2})\\.RDS$","\\1", 
                                           list.files(path = data_path, 
                                                      pattern = "\\d{4}-\\d{2}-\\d{2}\\.RDS$")), 
                                      tz = "America/New_York"),
                           stringsAsFactors = FALSE)
most_recent_datasets = paste(data_path, IVSA_datasets[which(IVSA_datasets$date == max(IVSA_datasets$date)),"files"], sep = "")

# IVSA
IVSA_raw_file = most_recent_datasets[grep("IVSA_raw", most_recent_datasets)]
IVSA_metadata_file = most_recent_datasets[grep("IVSA_metadata", most_recent_datasets)]
IVSA_raw = readRDS(IVSA_raw_file)
row.names(IVSA_raw) = IVSA_raw$DE_Subject
IVSA_metadata = readRDS(IVSA_metadata_file)
head(IVSA_raw)
```
Getting founders key file.

```{r founders_key_import, warning=FALSE}
founders_key = readRDS("C:/Users/dattau/Documents/R_IVSA_Troy/IVSA/IVSA/csna_project_4_behavior/data/cc_do_founders_key.RDS")
row.names(founders_key) = founders_key$strain
founders_key[9,1] = "J:DO"
founders_key[9,2] = "DO"
founders_key[9,5] = "#000000"
founders_key[9,6] = "#000000"
row.names(founders_key) = founders_key$strain
head(IVSA_raw)
```

### Filtering IVSA data for DO Founders

Getting IVSA data for the eight founders, old DO and CC  strains.

```{r filter_for_founders, warning=FALSE}
founders_IVSA = IVSA_raw %>%
  filter(DE_Strain %in% founders_key$strain)
#rm(list = c("IVSA_raw"))
founders_IVSA = as.data.frame(founders_IVSA)

founders_IVSA$DE_Strain = factor(founders_IVSA$DE_Strain,
                                 levels = founders_key$strain,
                                 ordered = TRUE)


Old_DO <- founders_IVSA %>% 
  filter(DE_birth_date_year == 2016,
         DE_Strain == "J:DO")


New_DO <- founders_IVSA %>% 
  filter(DE_birth_date_year != 2016,
         DE_Strain == "J:DO")

No_DO <- founders_IVSA %>% 
  filter(DE_Strain != "J:DO")

founders_IVSA <- merge(Old_DO,No_DO, all = TRUE)

ids <- c("A2063", "A2810", "A2418", "A3262", "A3261", "A3052", "A3257", "A3256", "A3051", "A2986")
rm.ids <- which(founders_IVSA$DE_Subject %in% ids)
founders_IVSA.p1  <- founders_IVSA[-rm.ids,]

founders_IVSA = founders_IVSA.p1

```

Getting CC key.

```{r}
CC_key = unique(IVSA_raw$DE_Strain)
CC_key = CC_key[grep("^CC\\d{3}/",CC_key)]
#CC_key = stringr::str_extract(CC_key, "^.{5}")
CC_key
CC_founders_IVSA_CC = IVSA_raw %>%
  filter(DE_Strain %in% c(CC_key))

CC_founders_IVSA_CC$DE_Strain = stringr::str_extract(CC_founders_IVSA_CC$DE_Strain, "^.{5}")

CC_key_abv <- unique(CC_founders_IVSA_CC$DE_Strain)

head(CC_founders_IVSA_CC)
```
Merging old DO, founders and CC. Generating a combined key.
```{r}
CC_founders_IVSA <- rbind(founders_IVSA, CC_founders_IVSA_CC)
Combined_key_abv <- c(founders_key$strain,CC_key_abv)
```

Merging old DO, founders and CC
```{r}
CC_founders_IVSA <- rbind(founders_IVSA, CC_founders_IVSA_CC)
```

The environment and datasets are ready for analysis.
# Analysis
### Getting IVSA extinction relevant columns
Selecting the relevant columns for Extinction relevant trait calculations. Choosing the following variables:
'EX_ALP_Total_s01'to 'EX_ALP_Total_s10': Active lever press for extinction sessions
'EX_ILP_Total_s01'to 'EX_ILP_Total_s10': Inactive lever press for extinction sessions
Changing the format from wide to long format.
```{r extinction_plot, warning=FALSE}
CC_extinction_df = CC_founders_IVSA %>%
  select(DE_Subject,
         DE_Strain,
         DE_Sex,
         EX_ALP_Total_s01,
         EX_ALP_Total_s02,
         EX_ALP_Total_s03,
         EX_ALP_Total_s04,
         EX_ALP_Total_s05,
         EX_ALP_Total_s06,
         EX_ALP_Total_s07,
         EX_ALP_Total_s08,
         EX_ALP_Total_s09,
         EX_ALP_Total_s10,
         EX_ILP_Total_s01,
         EX_ILP_Total_s02,
         EX_ILP_Total_s03,
         EX_ILP_Total_s04,
         EX_ILP_Total_s05,
         EX_ILP_Total_s06,
         EX_ILP_Total_s07,
         EX_ILP_Total_s08,
         EX_ILP_Total_s09,
         EX_ILP_Total_s10) %>%
  gather(key = Lever_Session, value = lever_presses, -DE_Subject, -DE_Strain, -DE_Sex) %>%
  mutate(Lever = gsub("^EX_([AI])LP_Total_s(\\d{2})$","\\1",Lever_Session),
         Session = as.numeric(gsub("^EX_([AI])LP_Total_s(\\d{2})$","\\2",Lever_Session))) %>%
  filter(!is.na(lever_presses))
```

### Getting IVSA Reinstatement relevant columns

```{r}
CC_founders_RI_ALP_df = CC_founders_IVSA %>%
  select(DE_Subject,
         DE_Strain,
         DE_Sex,
         RI_ALP_Total_s01,
         RI_ALP_Total_s02,
         RI_ALP_Total_s03) %>%
  gather(key = Session, value = ALP, -DE_Subject, -DE_Strain, -DE_Sex) %>%
  mutate(Session = as.numeric(gsub("^RI_ALP_Total_s(\\d{2})$","\\1",Session))) %>%
  filter(!is.na(ALP))
CC_founders_RI_ALP_df = as.data.frame(CC_founders_RI_ALP_df)

CC_founders_final_day_EX_ALP = CC_extinction_df %>%
  group_by(DE_Subject) %>%
  filter(!is.na(lever_presses) & Lever == "A") %>%
  summarize(ALP = lever_presses[which(Session == max(Session))])
CC_founders_final_day_EX_ALP = as.data.frame(CC_founders_final_day_EX_ALP)
row.names(CC_founders_final_day_EX_ALP) = CC_founders_final_day_EX_ALP$DE_Subject 

CC_founders_RI_ALP_df$final_EX_ALP = CC_founders_final_day_EX_ALP[CC_founders_RI_ALP_df$DE_Subject,"ALP"]
CC_founders_RI_ALP_df = CC_founders_RI_ALP_df[which(CC_founders_RI_ALP_df$Session == 1),]
CC_founders_RI_ALP_df$percent_final_EX = ((CC_founders_RI_ALP_df$ALP - CC_founders_RI_ALP_df$final_EX_ALP + 1) / (CC_founders_RI_ALP_df$final_EX_ALP + 1)) * 100
CC_founders_RI_ALP_df$ratio_final_EX = (CC_founders_RI_ALP_df$ALP + 1) / (CC_founders_RI_ALP_df$final_EX_ALP + 1)
CC_founders_RI_ALP_df
```
Looking for strain differences among founder and CC strains. No DO.
```{r}
 CC_RI_ALP_df <- CC_founders_RI_ALP_df %>%
  filter(DE_Strain != "J:DO")%>% 
  filter(DE_Strain %in% c(Combined_key_abv))
CC_RI_model <- anova(lm(percent_final_EX ~ DE_Strain, data = CC_RI_ALP_df))
CC_RI_model
CC_RI_model <- anova(lm(percent_final_EX ~ DE_Strain*DE_Sex, data = CC_RI_ALP_df))
CC_RI_model
```


Looking for strain differences among founder strains. No DO.

```{r}
Founders_RI_ALP_df <- CC_founders_RI_ALP_df %>%
  filter(DE_Strain != "J:DO") 
Founders_RI_model <- anova(lm(percent_final_EX ~ DE_Strain, data = Founders_RI_ALP_df))
Founders_RI_model
Founders_RI_model <- anova(lm(percent_final_EX ~ DE_Strain*DE_Sex, data = Founders_RI_ALP_df))
Founders_RI_model
```
Looking for strain differences among CC strains.

```{r}
 CC_RI_ALP_df <- CC_founders_RI_ALP_df %>%
  filter(DE_Strain %in% c(CC_key_abv))
CC_RI_model <- anova(lm(percent_final_EX ~ DE_Strain, data = CC_RI_ALP_df))
CC_RI_model
CC_RI_model <- anova(lm(percent_final_EX ~ DE_Strain*DE_Sex, data = CC_RI_ALP_df))
CC_RI_model
```

### IVSA Reinstatement plot
```{r}
CC_founders_RI_ALP_df_summary = CC_founders_RI_ALP_df %>%
  group_by(DE_Strain) %>%
  summarize(mean = mean(percent_final_EX, na.rm = TRUE),
            sem = sd(percent_final_EX, na.rm = TRUE) / sqrt(length(which(!is.na(percent_final_EX)))),
            n = length(which(!is.na(percent_final_EX)))) %>%
  mutate(ul = mean + sem,
         ll = mean - sem)
CC_founders_RI_ALP_df_summary= as.data.frame(CC_founders_RI_ALP_df_summary)
CC_founders_RI_ALP_df_summary$Color = rep("#CCCCCC", times = nrow(CC_founders_RI_ALP_df_summary))
row.names(CC_founders_RI_ALP_df_summary) = CC_founders_RI_ALP_df_summary$DE_Strain
CC_founders_RI_ALP_df_summary[row.names(founders_key),"Color"] = founders_key$collaborative_cross_color_broman
CC_founders_RI_ALP_df_summary  = CC_founders_RI_ALP_df_summary[which(!is.na(CC_founders_RI_ALP_df_summary$DE_Strain)),]
CC_founders_RI_ALP_df_summary$Label = paste(CC_founders_RI_ALP_df_summary$DE_Strain,
                                              sep="")
CC_founders_RI_ALP_df_summary$DE_Strain = factor(CC_founders_RI_ALP_df_summary$DE_Strain,
                                                   levels = CC_founders_RI_ALP_df_summary[order(CC_founders_RI_ALP_df_summary$mean),"DE_Strain"],
                                                   ordered = TRUE)

CC_founders_RI_ALP_plot = ggplot(data = CC_founders_RI_ALP_df_summary, aes(x = DE_Strain, y = mean, fill = DE_Strain)) +
  geom_col(color = "#000000") +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.5) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5), legend.position="none") +
  scale_fill_manual(values = CC_founders_RI_ALP_df_summary[order(CC_founders_RI_ALP_df_summary$DE_Strain),"Color"]) +
  xlab(NULL) +
  ylab("% inc. in Active Lever Presses during RI") +
  geom_text(aes(
    label= paste(as.character(),CC_founders_RI_ALP_df_summary$n, sep = ""),
    y= -1000), size = 2.75, hjust = 0.3, vjust = 0.3, angle = 90) 
CC_founders_RI_ALP_plot <- CC_founders_RI_ALP_plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
CC_founders_RI_ALP_plot
ggsave("CC_founders_RI_ALP_plot.pdf")
```
```{r}
CC_founders_RI_ALP_df_summary = CC_founders_RI_ALP_df %>%
  group_by(DE_Strain) %>%
  summarize(mean = mean(percent_final_EX, na.rm = TRUE),
            sem = sd(percent_final_EX, na.rm = TRUE) / sqrt(length(which(!is.na(percent_final_EX)))),
            n = length(which(!is.na(percent_final_EX)))) %>%
  mutate(ul = mean + sem,
         ll = mean - sem)
CC_founders_RI_ALP_df_summary= as.data.frame(CC_founders_RI_ALP_df_summary)
CC_founders_RI_ALP_df_summary$Color = rep("#CCCCCC", times = nrow(CC_founders_RI_ALP_df_summary))
row.names(CC_founders_RI_ALP_df_summary) = CC_founders_RI_ALP_df_summary$DE_Strain
CC_founders_RI_ALP_df_summary[row.names(founders_key),"Color"] = founders_key$collaborative_cross_color_broman
CC_founders_RI_ALP_df_summary  = CC_founders_RI_ALP_df_summary[which(!is.na(CC_founders_RI_ALP_df_summary$DE_Strain)),]
CC_founders_RI_ALP_df_summary$Label = paste(CC_founders_RI_ALP_df_summary$DE_Strain,
                                              sep="")
CC_founders_RI_ALP_df_summary$DE_Strain = factor(CC_founders_RI_ALP_df_summary$DE_Strain,
                                                   levels = CC_founders_RI_ALP_df_summary[order(CC_founders_RI_ALP_df_summary$mean),"DE_Strain"],
                                                   ordered = TRUE)
CC_founders_RI_ALP_df_summary
```

```{r}
CC_founders_RI_ALP_df_summary_no_CC035 <- CC_founders_RI_ALP_df_summary[which(CC_founders_RI_ALP_df_summary$DE_Strain != "CC035"),]
CC_founders_RI_ALP_plot_no_CC035 = ggplot(data = CC_founders_RI_ALP_df_summary_no_CC035, aes(x = DE_Strain, y = mean, fill = DE_Strain)) +
  geom_col(color = "#000000") +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.5) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position="none") +
  scale_fill_manual(values = CC_founders_RI_ALP_df_summary_no_CC035[order(CC_founders_RI_ALP_df_summary_no_CC035$DE_Strain),"Color"]) +
  xlab(NULL) +
  ylab("% inc. in Active Lever Presses during RI") +
 geom_text(aes(
    label= paste(as.character(),CC_founders_RI_ALP_df_summary_no_CC035$n, sep = ""),
    y= -150), size = 2.75, hjust = 0.3, vjust = 0.3, angle = 90) 
CC_founders_RI_ALP_plot_no_CC035 <- CC_founders_RI_ALP_plot_no_CC035 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))

CC_founders_RI_ALP_plot_no_CC035
ggsave("CC_founders_RI_ALP_plot_no_CC035.pdf")  

```



```{r}
CC_RI_ALP_df = CC_founders_IVSA %>%
  select(DE_Subject,
         DE_Strain,
         DE_Sex,
         RI_ALP_Total_s01,
         RI_ALP_Total_s02,
         RI_ALP_Total_s03) %>%
  gather(key = Session, value = ALP, -DE_Subject, -DE_Strain, -DE_Sex) %>%
  mutate(Session = as.numeric(gsub("^RI_ALP_Total_s(\\d{2})$","\\1",Session))) %>%
  filter(!is.na(ALP))
CC_RI_ALP_df = as.data.frame(CC_RI_ALP_df)

CC_final_day_EX_ALP = CC_extinction_df %>%
  group_by(DE_Subject) %>%
  filter(!is.na(lever_presses) & Lever == "A") %>%
  summarize(ALP = lever_presses[which(Session == max(Session))])
CC_final_day_EX_ALP = as.data.frame(CC_final_day_EX_ALP)
row.names(CC_final_day_EX_ALP) = CC_final_day_EX_ALP$DE_Subject 

CC_RI_ALP_df$final_EX_ALP = CC_final_day_EX_ALP[CC_RI_ALP_df$DE_Subject,"ALP"]
CC_RI_ALP_df = CC_RI_ALP_df[which(CC_RI_ALP_df$Session == 1),]
CC_RI_ALP_df$percent_final_EX = ((CC_RI_ALP_df$ALP - CC_RI_ALP_df$final_EX_ALP + 1) / (CC_RI_ALP_df$final_EX_ALP + 1)) * 100
CC_RI_ALP_df$ratio_final_EX = (CC_RI_ALP_df$ALP + 1) / (CC_RI_ALP_df$final_EX_ALP + 1)


CC_RI_ALP_summary = CC_RI_ALP_df %>%
  group_by(DE_Strain) %>%
  summarize(mean = mean(percent_final_EX, na.rm = TRUE),
            sem = sd(percent_final_EX, na.rm = TRUE) / sqrt(length(which(!is.na(percent_final_EX)))),
            n = length(which(!is.na(percent_final_EX)))) %>%
  mutate(ul = mean + sem, 
         ll = mean - sem,
         off_scale = ifelse(mean > 500, "YES", "no"))
CC_RI_ALP_summary = as.data.frame(CC_RI_ALP_summary)
row.names(CC_RI_ALP_summary) = CC_RI_ALP_summary$DE_Strain
CC_RI_ALP_summary$Color = rep("#CCCCCC", times = nrow(CC_RI_ALP_summary))
CC_RI_ALP_summary[row.names(founders_key),"Color"] = founders_key$collaborative_cross_color_broman
CC_RI_ALP_summary = CC_RI_ALP_summary[which(!is.na(CC_RI_ALP_summary$DE_Strain)),]
CC_RI_ALP_summary = CC_RI_ALP_summary[which(!is.na(CC_RI_ALP_summary$off_scale) |
                                              CC_RI_ALP_summary$off_scale == "YES"),]

CC_RI_ALP_summary$DE_Strain = factor(CC_RI_ALP_summary$DE_Strain,
                                     levels = CC_RI_ALP_summary[order(CC_RI_ALP_summary$mean),"DE_Strain"],
                                     ordered = TRUE)
RI_plot = ggplot(data = CC_RI_ALP_summary, aes(x = DE_Strain, y = mean, fill = DE_Strain)) +
  geom_col(color = "#000000") +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.2) +
  facet_wrap(vars(off_scale), scales = "free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5), legend.position="none",
        strip.background = element_blank(), strip.text.x = element_blank()) +
  scale_fill_manual(values = CC_RI_ALP_summary[order(CC_RI_ALP_summary$DE_Strain),"Color"]) +
  xlab(NULL) +
  ylab("Percent Increase in ALPs in Reinstatement Session 1")
RI_plot_grob = ggplotGrob(RI_plot)
facet.columns = RI_plot_grob$layout$l[grepl("panel", RI_plot_grob$layout$name)]
x.var = sapply(ggplot_build(RI_plot)$layout$panel_scales_x,
                function(l) length(l$range$range))
RI_plot_grob$widths[facet.columns] = RI_plot_grob$widths[facet.columns] * x.var
RI_plot

RI_plot<- RI_plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))

RI_plot
ggsave("RI_plot.pdf") 
```
```{r}
DO_founders_RI_histogram <- CC_founders_RI_ALP_df %>%
  filter(DE_Strain %in% founders_key$strain)  

DO_founders_RI_histogram %>%  ggplot(aes(x = percent_final_EX, y = DE_Strain, height = stat(density), fill = DE_Strain)) + 
  geom_density_ridges(
    stat = "binline", bins = 30, scale = 0.95,
    draw_baseline = FALSE) + 
  scale_fill_manual(values = founders_key[which(founders_key$strain %in% DO_founders_RI_histogram$DE_Strain),"collaborative_cross_color_broman"]) + 
  theme_bw() +
  theme(panel.grid = element_blank())+ theme(legend.position = "none")+   theme(aspect.ratio = 1)

```

```{r}
DO_founders_RI_histogram <- CC_founders_RI_ALP_df %>%
  filter(DE_Strain %in% founders_key$strain)  

DO_founders_RI_histogram <- DO_founders_RI_histogram[which(DO_founders_RI_histogram$percent_final_EX <=5000), ]

DO_founders_RI_histogram %>%  ggplot(aes(x = percent_final_EX, y = DE_Strain, height = stat(density), fill = DE_Strain)) + 
  geom_density_ridges(
    stat = "binline", bins = 30, scale = 0.95,
    draw_baseline = FALSE) + 
  scale_fill_manual(values = founders_key[which(founders_key$strain %in% DO_founders_RI_histogram$DE_Strain),"collaborative_cross_color_broman"]) + 
  theme_bw() +
  theme(panel.grid = element_blank())+ theme(legend.position = "none")+   theme(aspect.ratio = 1)

```

## Reinstatement plot for founders and DO.
Last day of cocaine self-administration, last day of extinction and two reinstatement sessions. Adjusted y axis. Unadjusted y-xis.
```{r}
founders_reinstatement_df = founders_IVSA %>%
  dplyr::select(DE_Subject,
         DE_Strain,
         DE_Sex,
         DR_ALP_TfASoTD_1p8mgkg,
         EX_ALP_Total_s07,
         RI_ALP_Total_s01,
         RI_ALP_Total_s02,
         RI_ALP_Total_s03,
         DR_ILP_TfASoTD_1p8mgkg,
         EX_ILP_Total_s07,
         RI_ILP_Total_s01,
         RI_ILP_Total_s02,
         RI_ILP_Total_s03) 

  colnames(founders_reinstatement_df)[4] <- "RI_ALP_Total_s00"
  colnames(founders_reinstatement_df)[5] <- "RI_ALP_Total_s07"
  colnames(founders_reinstatement_df)[9] <- "RI_ILP_Total_s00"
  colnames(founders_reinstatement_df)[10] <- "RI_ILP_Total_s07"

  founders_reinstatement_df <- founders_reinstatement_df%>%
  gather(key = Lever_Session, value = lever_presses, -DE_Subject, -DE_Strain, -DE_Sex) %>%
  mutate(Lever = gsub("^RI_([AI])LP_Total_s(\\d{2})$","\\1",Lever_Session),
         Session = as.numeric(gsub("^RI_([AI])LP_Total_s(\\d{2})$","\\2",Lever_Session))) %>%
  filter(!is.na(lever_presses), Session < 8)

founders_reinstatement_summary = founders_reinstatement_df %>%
  group_by(DE_Strain, Lever, Session) %>%
  summarize(mean = mean(lever_presses, na.rm = TRUE),
            n = length(which(!is.na(lever_presses))),
            sem = sd(lever_presses, na.rm = TRUE) / sqrt(length(which(!is.na(lever_presses))))) %>%
  mutate(ul = mean + sem, ll = mean - sem)

founders_reinstatement_summary$Session <- gsub(founders_reinstatement_summary$Session, pattern = "2", replacement = "3")
founders_reinstatement_summary$Session <- gsub(founders_reinstatement_summary$Session, pattern = "1", replacement = "2")
founders_reinstatement_summary$Session <- gsub(founders_reinstatement_summary$Session, pattern = "7", replacement = "1")

founders_reinstatement_summary$Session <- as.numeric(founders_reinstatement_summary$Session)

reinstatment_plot = ggplot(data = founders_reinstatement_summary, aes(x = Session, 
                                               y = mean, 
                                               color = DE_Strain,
                                               linetype = Lever,
                                               shape = Lever,
                                               fill = Lever)) +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.2) +
  geom_line(size=1.05) +
  geom_point() +
  theme_bw() +
  facet_grid(cols = vars(DE_Strain)) +
  scale_x_continuous(limits = c(0,3), breaks = c(0,1,2,3), labels = c("DR-1.8mgkg","EX-LastSession","RI-S01","RI-S02")) +
  scale_color_manual(values = founders_key[c("A/J","C57BL/6J","NOD/ShiLtJ","NZO/HlLtJ","CAST/EiJ","PWK/PhJ","WSB/EiJ", "J:DO"),"collaborative_cross_color_broman"]) +
  scale_fill_manual(values = c("#000000","#FFFFFF")) +
  scale_shape_manual(values = c(16,21)) +
  theme(panel.grid = element_blank()) +
  theme(text = element_text(size=18)) +
  scale_y_continuous(limits =  c(0,375)) +
  xlab("Session") +
  ylab("Number of Lever Presses") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


reinstatment_plot

reinstatment_plot_bar = ggplot(data = founders_reinstatement_summary) +
  geom_bar(aes(x = Session, y = mean, fill = Lever), stat = "identity", position = "dodge") + 
  #geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.2) +
  #geom_point() +
  theme_bw() +
  facet_grid(cols = vars(DE_Strain)) +
  scale_x_continuous(limits = c(0,3), breaks = c(0,1,2,3), labels = c("DR-1.8mgkg","EX-LastSession","RI-S01","RI-S02")) +
  scale_color_manual(values = founders_key[c("A/J","C57BL/6J","NOD/ShiLtJ","NZO/HlLtJ","CAST/EiJ","PWK/PhJ","WSB/EiJ","J:DO"),"collaborative_cross_color_broman"]) +
  #scale_fill_manual(values = c("#000000","#FFFFFF")) +
  #scale_shape_manual(values = c(16,21)) +
  theme(panel.grid = element_blank()) +
  theme(text = element_text(size=18)) +
  scale_y_continuous(limits =  c(0,375)) +
  xlab("Session") +
  ylab("Number of Lever Presses") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


reinstatment_plot = ggplot(data = founders_reinstatement_summary, aes(x = Session, 
                                               y = mean, 
                                               color = DE_Strain,
                                               linetype = Lever,
                                               shape = Lever,
                                               fill = Lever)) +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.2) +
  geom_line(size=1.05) +
  geom_point() +
  theme_bw() +
  facet_grid(cols = vars(DE_Strain)) +
  scale_x_continuous(limits = c(0,3), breaks = c(0,1,2,3), labels = c("DR-1.8mgkg","EX-LastSession","RI-S01","RI-S02")) +
  scale_color_manual(values = founders_key[c("A/J","C57BL/6J","NOD/ShiLtJ","NZO/HlLtJ","CAST/EiJ","PWK/PhJ","WSB/EiJ","J:DO"),"collaborative_cross_color_broman"]) +
  scale_fill_manual(values = c("#000000","#FFFFFF")) +
  scale_shape_manual(values = c(16,21)) +
  theme(panel.grid = element_blank()) +
  theme(text = element_text(size=18)) +
  #scale_y_continuous(limits =  c(0,375)) +
  xlab("Session") +
  ylab("Number of Lever Presses") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


reinstatment_plot
```
## Reinstatement plot for founders and DO.
Last day \of extinction and two reinstatement sessions. Adjusted y axis. Unadjusted y-xis.


```{r}
founders_reinstatement_df = founders_IVSA %>%
  dplyr::select(DE_Subject,
         DE_Strain,
         DE_Sex,
         EX_ALP_Total_s07,
         RI_ALP_Total_s01,
         RI_ALP_Total_s02,
         RI_ALP_Total_s03,
         EX_ILP_Total_s07,
         RI_ILP_Total_s01,
         RI_ILP_Total_s02,
         RI_ILP_Total_s03) 

  colnames(founders_reinstatement_df)[4] <- "RI_ALP_Total_s00"

  colnames(founders_reinstatement_df)[8] <- "RI_ILP_Total_s00"

  founders_reinstatement_df <- founders_reinstatement_df%>%
  gather(key = Lever_Session, value = lever_presses, -DE_Subject, -DE_Strain, -DE_Sex) %>%
  mutate(Lever = gsub("^RI_([AI])LP_Total_s(\\d{2})$","\\1",Lever_Session),
         Session = as.numeric(gsub("^RI_([AI])LP_Total_s(\\d{2})$","\\2",Lever_Session))) %>%
  filter(!is.na(lever_presses), Session < 8)

founders_reinstatement_summary = founders_reinstatement_df %>%
  group_by(DE_Strain, Lever, Session) %>%
  summarize(mean = mean(lever_presses, na.rm = TRUE),
            n = length(which(!is.na(lever_presses))),
            sem = sd(lever_presses, na.rm = TRUE) / sqrt(length(which(!is.na(lever_presses))))) %>%
  mutate(ul = mean + sem, ll = mean - sem)



founders_reinstatement_summary$Session <- as.numeric(founders_reinstatement_summary$Session)

reinstatment_plot = ggplot(data = founders_reinstatement_summary, aes(x = Session, 
                                               y = mean, 
                                               color = DE_Strain,
                                               linetype = Lever,
                                               shape = Lever,
                                               fill = Lever)) +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.2) +
  geom_line(size=1.05) +
  geom_point() +
  theme_bw() +
  facet_grid(cols = vars(DE_Strain)) +
  scale_x_continuous(limits = c(0,2), breaks = c(0,1,2), labels = c("EX-LastSession","RI-S01","RI-S02")) +
  scale_color_manual(values = founders_key[c("A/J","C57BL/6J","NOD/ShiLtJ","NZO/HlLtJ","CAST/EiJ","PWK/PhJ","WSB/EiJ", "J:DO"),"collaborative_cross_color_broman"]) +
  scale_fill_manual(values = c("#000000","#FFFFFF")) +
  scale_shape_manual(values = c(16,21)) +
  theme(panel.grid = element_blank()) +
  theme(text = element_text(size=18)) +
  scale_y_continuous(limits =  c(0,375)) +
  xlab("Session") +
  ylab("Number of Lever Presses") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


reinstatment_plot

reinstatment_plot_bar = ggplot(data = founders_reinstatement_summary) +
  geom_bar(aes(x = Session, y = mean, fill = Lever), stat = "identity", position = "dodge") + 
  #geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.2) +
  #geom_point() +
  theme_bw() +
  facet_grid(cols = vars(DE_Strain)) +
  scale_x_continuous(limits = c(0,2), breaks = c(0,1,2), labels = c("EX-LastSession","RI-S01","RI-S02")) +
  scale_color_manual(values = founders_key[c("A/J","C57BL/6J","NOD/ShiLtJ","NZO/HlLtJ","CAST/EiJ","PWK/PhJ","WSB/EiJ","J:DO"),"collaborative_cross_color_broman"]) +
  #scale_fill_manual(values = c("#000000","#FFFFFF")) +
  #scale_shape_manual(values = c(16,21)) +
  theme(panel.grid = element_blank()) +
  theme(text = element_text(size=18)) +
  scale_y_continuous(limits =  c(0,375)) +
  xlab("Session") +
  ylab("Number of Lever Presses") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


reinstatment_plot = ggplot(data = founders_reinstatement_summary, aes(x = Session, 
                                               y = mean, 
                                               color = DE_Strain,
                                               linetype = Lever,
                                               shape = Lever,
                                               fill = Lever)) +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.2) +
  geom_line(size=1.05) +
  geom_point() +
  theme_bw() +
  theme(legend.position="none")+
  facet_grid(cols = vars(DE_Strain)) +
  scale_x_continuous(limits = c(0,2), breaks = c(0,1,2), labels = c("EX-LastSession","RI-S01","RI-S02")) +
  scale_color_manual(values = founders_key[c("A/J","C57BL/6J","NOD/ShiLtJ","NZO/HlLtJ","CAST/EiJ","PWK/PhJ","WSB/EiJ","J:DO"),"collaborative_cross_color_broman"]) +
  scale_fill_manual(values = c("#000000","#FFFFFF")) +
  scale_shape_manual(values = c(16,21)) +
  theme(panel.grid = element_blank()) +
  theme(text = element_text(size=18)) +
  #scale_y_continuous(limits =  c(0,375)) +
  xlab("Session") +
  ylab("Number of Lever Presses") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


reinstatment_plot
ggsave("reinstatment_plot.pdf")
```

Selecting the relevant columns for Reinstatement relevant trait calculations. Choosing the following variables:
'EX_ALP_Total_*last session': Active lever press for the final extinction session
```{r extinction_plot, warning=FALSE}
founders_extinction_df = founders_IVSA %>%
  select(DE_Subject,
         DE_Strain,
         DE_Sex,
         EX_ALP_Total_s01,
         EX_ALP_Total_s02,
         EX_ALP_Total_s03,
         EX_ALP_Total_s04,
         EX_ALP_Total_s05,
         EX_ALP_Total_s06,
         EX_ALP_Total_s07) 
founders_extinction_df <- founders_extinction_df        %>%
  gather(key = Lever_Session, value = lever_presses, -DE_Subject, -DE_Strain, -DE_Sex) %>%
  mutate(Lever = gsub("^EX_([AI])LP_Total_s(\\d{2})$","\\1",Lever_Session),
         Session = as.numeric(gsub("^EX_([AI])LP_Total_s(\\d{2})$","\\2",Lever_Session))) %>%
  filter(!is.na(lever_presses), Session <= 8)

  founders_extinction_df 
```

'RI_Inf_Total_*s01_03'three extinction sessions.
Calculating %inc. in active lever press in first reinstatement sessions relative to last extinction session as a percent and also as a ratio.
Changing the format from wide to long format.

```{r}

Founders_RI_df = founders_IVSA %>%
  select(DE_Subject,
         DE_Strain,
         DE_Sex,
         RI_Inf_Total_s01,
         RI_Inf_Total_s02,
         RI_Inf_Total_s03) %>%
  gather(key = Session, value = Infusions, -DE_Subject, -DE_Strain, -DE_Sex) %>%
  mutate(Session = as.numeric(gsub("^RI_Inf_Total_s(\\d{2})$","\\1",Session))) %>%
  filter(!is.na(Infusions))

Founders_RI_ALP_df = founders_IVSA %>%
  select(DE_Subject,
         DE_Strain,
         DE_Sex,
         RI_ALP_Total_s01,
         RI_ALP_Total_s02,
         RI_ALP_Total_s03) %>%
  gather(key = Session, value = ALP, -DE_Subject, -DE_Strain, -DE_Sex) %>%
  mutate(Session = as.numeric(gsub("^RI_ALP_Total_s(\\d{2})$","\\1",Session))) %>%
  filter(!is.na(ALP))
Founders_RI_ALP_df = as.data.frame(Founders_RI_ALP_df)
Founders_RI_ALP_df
Founders_final_day_EX_ALP = founders_extinction_df %>%
  group_by(DE_Subject) %>%
  filter(!is.na(lever_presses) & Lever == "A") %>%
  summarize(ALP = lever_presses[which(Session == max(Session))])
Founders_final_day_EX_ALP = as.data.frame(Founders_final_day_EX_ALP)
row.names(Founders_final_day_EX_ALP) = Founders_final_day_EX_ALP$DE_Subject 
Founders_final_day_EX_ALP 
Founders_RI_ALP_df
Founders_RI_ALP_df$final_EX_ALP = Founders_final_day_EX_ALP[Founders_RI_ALP_df$DE_Subject,"ALP"]
Founders_RI_ALP_df
Founders_RI_ALP_df = Founders_RI_ALP_df[which(Founders_RI_ALP_df$Session == 1),]
Founders_RI_ALP_df$percent_inc_final_EX = ((Founders_RI_ALP_df$ALP - Founders_RI_ALP_df$final_EX_ALP + 1) / (Founders_RI_ALP_df$final_EX_ALP + 1)) * 100
Founders_RI_ALP_df$ratio_final_EX = (Founders_RI_ALP_df$ALP + 1) / (Founders_RI_ALP_df$final_EX_ALP + 1)
Founders_RI_ALP_df
```
Excluding DO. Asessing Strain and Sex effects on %inc in lever presses on first reinstatement session relative to extinction.
```{r}
Founders_RI_ALP_df <- Founders_RI_ALP_df %>%
  filter(DE_Strain != "J:DO") %>%
  filter(DE_Strain %in% founders_key$strain)
anova(lm(percent_inc_final_EX ~ DE_Strain * DE_Sex, data = Founders_RI_ALP_df))
anova(lm(ratio_final_EX ~ DE_Strain * DE_Sex, data = Founders_RI_ALP_df))
anova(lm(percent_inc_final_EX ~ DE_Strain, data = Founders_RI_ALP_df))
anova(lm(ratio_final_EX ~ DE_Strain, data = Founders_RI_ALP_df))

continuous.model <- lm(ALP~ DE_Strain, data = Founders_RI_ALP_df)
summary(continuous.model)
plot(continuous.model)
coef(continuous.model)
```
Asessing strain differences on increase in lever press during first reinstatement relative to last session of extinction as repeated measures. 

```{r}
Founders_RI_ALP_df <- Founders_RI_ALP_df %>%
  filter(DE_Strain != "J:DO") %>%
  filter(DE_Strain %in% founders_key$strain)

Founders_RI_ALP_df %>%
  filter(!is.na(percent_inc_final_EX)) %>%
  group_by(DE_Strain) %>%
  summarize(n = length(percent_inc_final_EX))


Founders_RI_ALP_df = Founders_RI_ALP_df %>%
  rename(
    "2" = ALP, 
    "1" = final_EX_ALP
    )
Founders_RI_ALP_df 

Founders_RI_ALP_df  = Founders_RI_ALP_df %>%
  gather(key = Session, value = ALP, "2":"1", -DE_Subject, -DE_Strain, -DE_Sex) 
Founders_RI_ALP_df 

Founders_RI_ALP_df  = as.data.frame(Founders_RI_ALP_df)

Founders_RI_ALP_df$Session <- as.factor(Founders_RI_ALP_df$Session)   


complete_obs <- na.omit(Founders_RI_ALP_df)
complete_obs

complete_obs %>%
  group_by(DE_Strain)%>%
  summarize(n())

complete_obs$Session <- as.factor(complete_obs$Session)
model3 =ezANOVA(
  data = complete_obs,
  dv = ALP,
  wid = DE_Subject,
  within = .(Session),
  between = .(DE_Strain),
  #between = .(DE_Strain, DE_Sex),
  detailed = TRUE,
  type = 3
)
print(model3)
summary(model3)

```
Excluding DO. Asessing Strain and Sex effects on %inc in percent active lever presses on reinstatement relative to extinction and ratio of pALP on extinction and reinstatement.
```{r}
Founders_RI_pALP_df = founders_IVSA %>%
  select(DE_Subject,
         DE_Strain,
         DE_Sex,
         RI_pAP_Total_s01,
         RI_pAP_Total_s02,
         RI_pAP_Total_s03) %>%
  gather(key = Session, value = pALP, -DE_Subject, -DE_Strain, -DE_Sex) %>%
  mutate(Session = as.numeric(gsub("^RI_pAP_Total_s(\\d{2})$","\\1",Session))) %>%
  filter(!is.na(pALP))
Founders_RI_pALP_df = as.data.frame(Founders_RI_pALP_df)
Founders_RI_pALP_df
Founders_final_day_EX_pALP = founders_extinction_df %>%
  group_by(DE_Subject) %>%
  filter(!is.na(lever_presses) & Lever == "A") %>%
  summarize(pALP = lever_presses[which(Session == max(Session))])
Founders_final_day_EX_pALP = as.data.frame(Founders_final_day_EX_pALP)
row.names(Founders_final_day_EX_pALP) = Founders_final_day_EX_pALP$DE_Subject 
Founders_final_day_EX_pALP 
Founders_RI_pALP_df
Founders_RI_pALP_df$final_EX_pALP = Founders_final_day_EX_pALP[Founders_RI_pALP_df$DE_Subject,"pALP"]
Founders_RI_pALP_df
Founders_RI_pALP_df = Founders_RI_pALP_df[which(Founders_RI_pALP_df$Session == 1),]
Founders_RI_pALP_df$percent_inc_final_EX = ((Founders_RI_pALP_df$pALP - Founders_RI_pALP_df$final_EX_pALP + 1) / (Founders_RI_pALP_df$final_EX_pALP + 1)) * 100
Founders_RI_pALP_df$ratio_final_EX = (Founders_RI_pALP_df$pALP + 1) / (Founders_RI_pALP_df$final_EX_pALP + 1)
Founders_RI_pALP_df <- Founders_RI_pALP_df %>%
  filter(DE_Strain != "J:DO") %>%
  filter(DE_Strain %in% founders_key$strain)
Founders_RI_pALP_df
```

```{r}

Founders_RI_pALP_df %>%
  filter(!is.na(percent_inc_final_EX)) %>%
  group_by(DE_Strain) %>%
  summarize(n = length(percent_inc_final_EX))


Founders_RI_pALP_df = Founders_RI_pALP_df %>%
  rename(
    "2" = pALP, 
    "1" = final_EX_pALP
    )
Founders_RI_ALP_df 

Founders_RI_pALP_df  = Founders_RI_pALP_df %>%
  gather(key = Session, value = pALP, "2":"1", -DE_Subject, -DE_Strain, -DE_Sex) 
Founders_RI_pALP_df 

Founders_RI_pALP_df  = as.data.frame(Founders_RI_pALP_df)

Founders_RI_pALP_df$Session <- as.factor(Founders_RI_pALP_df$Session)   


complete_obs <- na.omit(Founders_RI_pALP_df)
complete_obs

complete_obs %>%
  group_by(DE_Strain)%>%
  summarize(n())

complete_obs$Session <- as.factor(complete_obs$Session)
model3 =ezANOVA(
  data = complete_obs,
  dv = pALP,
  wid = DE_Subject,
  within = .(Session),
  between = .(DE_Strain),
  #between = .(DE_Strain, DE_Sex),
  detailed = TRUE,
  type = 3
)
print(model3)

```
Preparing dataframe to asess strain differences on increase in percent active lever press during reinstatement as an average of session 1 and 2 relative to last session of extinction as repeated measures. 
```{r}
Founders_RI_pALP_df = founders_IVSA %>%
  select(DE_Subject,
         DE_Strain,
         DE_Sex,
         RI_pAP_Total_s01,
         RI_pAP_Total_s02,
         RI_pAP_Total_s03) %>%
  gather(key = Session, value = pALP, -DE_Subject, -DE_Strain, -DE_Sex) %>%
  mutate(Session = as.numeric(gsub("^RI_pAP_Total_s(\\d{2})$","\\1",Session))) %>%
  filter(!is.na(pALP))
Founders_RI_pALP_df = as.data.frame(Founders_RI_pALP_df)
Founders_RI_pALP_df
Founders_final_day_EX_pALP = founders_extinction_df %>%
  group_by(DE_Subject) %>%
  filter(!is.na(lever_presses) & Lever == "A") %>%
  summarize(pALP = lever_presses[which(Session == max(Session))])
Founders_final_day_EX_pALP = as.data.frame(Founders_final_day_EX_pALP)
row.names(Founders_final_day_EX_pALP) = Founders_final_day_EX_pALP$DE_Subject 
Founders_final_day_EX_pALP 
Founders_RI_pALP_df
Founders_RI_pALP_df_wide <- spread(Founders_RI_pALP_df, Session, pALP)
Founders_RI_pALP_df_wide 
Founders_RI_pALP_df_wide $final_EX_pALP = Founders_final_day_EX_pALP[Founders_RI_pALP_df_wide$DE_Subject,"pALP"]
Founders_RI_pALP_df_wide
Founders_RI_pALP_df_wide = Founders_RI_pALP_df_wide %>%
  rename(
   RI_1 = "1", 
   RI_2 = "2"
    )
Founders_RI_pALP_df_wide$avg_RI_pALP = (Founders_RI_pALP_df_wide$RI_1 + Founders_RI_pALP_df_wide$RI_2) / 2
Founders_RI_pALP_df_wide <- Founders_RI_pALP_df_wide %>%
  filter(DE_Strain != "J:DO") %>%
  filter(DE_Strain %in% founders_key$strain)
Founders_RI_pALP_df_wide 
```


```{r}
Founders_RI_avg_pALP_df  = Founders_RI_pALP_df_wide %>%
  gather(key = Session, value = pALP, "final_EX_pALP":"avg_RI_pALP", -DE_Subject, -DE_Strain, -DE_Sex) 
Founders_RI_avg_pALP_df


Founders_RI_avg_pALP_df  = as.data.frame(Founders_RI_avg_pALP_df)

Founders_RI_avg_pALP_df$Session <- as.factor(Founders_RI_avg_pALP_df$Session)   


complete_obs <- na.omit(Founders_RI_avg_pALP_df)
complete_obs

complete_obs %>%
  group_by(DE_Strain)%>%
  summarize(n())

complete_obs$Session <- as.factor(complete_obs$Session)
model3 =ezANOVA(
  data = complete_obs,
  dv = pALP,
  wid = DE_Subject,
  within = .(Session),
  between = .(DE_Strain),
  #between = .(DE_Strain, DE_Sex),
  detailed = TRUE,
  type = 3
)
print(model3)

```