---
title: "Process_all_phenotypes_in_founder_CC_DO"
author: "Hao He"
output:
  workflowr::wflow_html:
    toc: true
    toc_float: true
    code_folding: "hide"
---

### Purpose of analysis
* 1, Summarize and plot phenotypes for RL, SENS, IVSA and CR across Founders, CC and DO
* 2, Genetic correlations in Founders and/or CC 
* 3, Process OFA, NPP, HB, LD, RL, SENS, IVSA and CR for DO and correlation
* 4, Generate a file containing all phenotype of above domains for qtl mapping

### 1, Working directory
All chunks will run in the working folder for this experiment.
```{r setup, warning = FALSE}
script_root_dir = "/projects/csna/csna_workflow"
```

### 2, R libraries
Calling R libraries necessary for this analysis.
```{r libraries, warning = FALSE}
library(gridExtra)
library(psych)
library(ggcorrplot)
library(ggplot2)
library(lubridate)
library(ggpubr)
library(lme4)
library(lmerTest)
library(ggrepel)
library(cowplot)
library(lsr)
library(pwr)
library(tidyverse)
# Suppress summarise info
options(dplyr.summarise.inform = FALSE)
options(stringsAsFactors = FALSE)
```

### 3, Utility functions
```{r Utility functions, warning = FALSE}
#founder key
founders_key = readRDS("data/cc_do_founders_key.rds")
row.names(founders_key) = founders_key$strain

#norm_rank_transform
norm_rank_transform = function(x, c = (0)) {
  stopifnot(is.numeric(x) & is.vector(x))
  x_noNA = which(!is.na(x))
  N = length(x_noNA)
  x[x_noNA] = qnorm((rank(x[x_noNA], ties.method = "random") - c) / (N - (2 * c) + 1))
  return(x)
}
# Getting a normal rank transformation function that can handle `NA` values from the following formula:
# 
# $Y^t_i={\phi}^{-1}({{r_i-c}\over{N-2c+1}})$
# 
# where $r_i$ is the rank, $N$ is the number of ranked samples, $\phi^{-1}$ is the inverse normal funcion (equivalent to `qnorm()` in R). The original recommended value of $c=3/8$ was per [Beasley and Erickson, 2010](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2921808/), but we use $c=0$. Our work appears to meet their sample size recommendations.

#powerTransform
powerTransform <- function(y, lambda1, lambda2 = NULL, method = "boxcox") {
  boxcoxTrans <- function(x, lam1, lam2 = NULL) {
    # if we set lambda2 to zero, it becomes the one parameter transformation
    lam2 <- ifelse(is.null(lam2), 0, lam2)
    if (lam1 == 0L) {
      log(y + lam2)
    } else {
      (((y + lam2)^lam1) - 1) / lam1
    }
  }
  switch(method
         , boxcox = boxcoxTrans(y, lambda1, lambda2)
         , tukey = y^lambda1
  )
}
#summary_func
summary_func <- list(
  mean = ~mean(.x, na.rm = TRUE),
  n    = ~length(which(!is.na(.x))),
  ul   = ~(mean(.x, na.rm = TRUE) + (sd(.x, na.rm = TRUE)/sqrt(length(which(!is.na(.x)))))),
  ll   = ~(mean(.x, na.rm = TRUE) - (sd(.x, na.rm = TRUE)/sqrt(length(which(!is.na(.x))))))
)
#summary_extremesCC_func
summary_extremesCC_func <- list(
  mean = ~MASS::huber(na.omit(.x))$mu,
  n    = ~length(which(!is.na(.x))),
  ul   = ~(MASS::huber(na.omit(.x))$mu + (sd(.x, na.rm = TRUE)/sqrt(length(which(!is.na(.x)))))),
  ll   = ~(MASS::huber(na.omit(.x))$mu - (sd(.x, na.rm = TRUE)/sqrt(length(which(!is.na(.x))))))
)
#outliers are outside of mean +/- 3SD
remove_outliers <- function(x, thres = 3, na.rm = TRUE, ...) {
  y <- x
  y[y <= mean(y,na.rm = na.rm) - thres * sd(y, na.rm = na.rm) | y >= mean(y,na.rm = na.rm) + thres * sd(y, na.rm = na.rm)] <- NA
  y
}
#rankz transform
rz.transform <- function(y){
  rankY=rank(y, ties.method="average", na.last="keep")
  rzT=qnorm(rankY/(length(na.exclude(rankY))+1))
  return(rzT)
}
#plot.summary
plot.summary <- function(summary.dat, pheno.list, out.pdf){
  p <- list()
  for(i in pheno.list){
    d <- summary.dat %>% 
      left_join(founders_key[,c(2,5)], by = c("Strain" = "collaborative_cross_abbreviation")) %>%
      rename(color = collaborative_cross_color_broman) %>%
      mutate(color = replace_na(color, "grey")) %>%
      mutate(color = case_when(
        Strain == "J:DO" ~ "springgreen",
        TRUE             ~ as.character(color))) %>%
      relocate(color, .after = Strain) %>%
      select(c("Strain", "color", contains(all_of(i)))) %>%
      rename_with(~(.x = c("Strain", "color", "mean", "n", "ul", "ll"))) %>%
      filter(n >= 1) %>%
      arrange(mean) %>%
      mutate(across(Strain, ~factor(.x, levels = as.character(.x)))) %>%
      mutate(xlabel = paste0(Strain," (n = ", n, ")"))
    p[[i]] <- ggplot(data = d, aes(x = Strain, y = mean, fill = Strain)) +
      geom_col(color = "#000000") +
      geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.5) +
      theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") +
      scale_fill_manual(values = d$color) +
      labs(x ="", y = paste0("Mean of ", i)) +
      scale_x_discrete(labels = d$xlabel) +
      scale_y_continuous(#expand = c(0,0),
                         limits = c(ifelse(min(c(d$ll, d$mean), na.rm = TRUE) > 0, 
                                           0, 
                                           min(c(d$ll, d$mean), na.rm = TRUE)+0.12*min(c(d$ll, d$mean), na.rm = TRUE)),
                                    max(c(d$ul, d$mean), na.rm = TRUE) + 0.12*max(c(d$ul, d$mean), na.rm = TRUE)))+
      theme_bw(base_size = 18) +
      theme(plot.title = element_text(hjust = 0.5),
            panel.background = element_blank(),
            panel.border = element_blank(),
            legend.position = "none",
            panel.grid.major.x = element_blank(),
            panel.grid.minor.x = element_blank(),
            axis.line = element_line(colour = "black"),
            axis.text.x = element_text(angle = 90, hjust = 0.95,vjust = 0.2))
    print(p[[i]])
  }
  ggsave(file = paste0(out.pdf,".pdf"), 
         plot = marrangeGrob(p, nrow = 1, ncol=1, top = NULL, bottom = NULL),
         device = "pdf", height = 8, width = 14)
}
#extremesCC_func
extremesCC_func <- function(cc.data, pheno.list){
  extremesCC <- list()
  for(i in pheno.list){
    #print(i)
    data  = cc.data %>%
      dplyr::filter(!is.na(.data[[i]]))
      #split data into three lists, male, female and bothsex.
      extremesCC[[i]] <- list(both   = data %>% mutate(Sex =  "Both"),
                              male   = data %>% filter(Sex == "Male"),
                              female = data %>% filter(Sex == "Female")) %>%
        map(function(x){
              y <- x %>%
                dplyr::select(c("Mouse.ID", "Strain", "Sex", contains(all_of(i)))) %>%
                dplyr::mutate(population_mean = mean(.[[all_of(i)]], na.rm = TRUE)) %>%
                group_by(Strain, Sex) %>% #Strain and Sex
                summarise(across(c(all_of(i), population_mean), summary_func)) %>%
                dplyr::select(1:7) %>%
                dplyr::mutate(across(everything(), ~replace_na(.x, NA))) %>%
                rename_with(~(.x = c("Strain", "Sex" ,"mean", "n", "ul", "ll", "pop_mean"))) %>%
                dplyr::mutate(diff_mean = mean - pop_mean)
              
              #t test for each strain
              dat.s.res <- list()
              for(s in unique(x$Strain)){
                #print(s)
                dat.s <- x %>% dplyr::filter(Strain == s)
                #print(dim(dat.s))
                v <- dat.s[[all_of(i)]]
                if(dim(dat.s)[[1]] >= 2 & length(unique(v)) != 1){
                  t      <- t.test(v, 
                                   mu = as.numeric(unique(y$pop_mean)), 
                                   alternative="two.sided", 
                                   conf.level=0.95)
                  t.stat <- t$statistic
                  t.p    <- t$p.value
                  lo.CI  <- t$conf.int[1]
                  up.CI  <- t$conf.int[2]
                  power  <- round(pwr.t.test(d = cohensD(v, 
                                                         mu = as.numeric(unique(y$pop_mean))),
                                             n = length(v),
                                             sig.level = 0.05,
                                             type = "one.sample",
                                             alternative = "two.sided")$power,2)
                  sample.size.at.0.8 <- ifelse(cohensD(v, 
                                                       mu = as.numeric(unique(y$pop_mean))) < 5,
                                               round(pwr.t.test(d = cohensD(v, 
                                                                            mu = as.numeric(unique(y$pop_mean))), 
                                                                sig.level = 0.05,
                                                                type="one.sample",
                                                                alternative="two.sided", 
                                                                power = 0.8)$n, 1), 
                                               round(pwr.t.test(d = 5, 
                                                                sig.level = 0.05,
                                                                type = "one.sample",
                                                                alternative = "two.sided", 
                                                                power = 0.8)$n, 1))
                  dat.s.res[[s]] <- data.frame(Strain = s,
                                               t.stat = t.stat, 
                                               t.p    = t.p, 
                                               lo.CI  = lo.CI, 
                                               up.CI  = up.CI, 
                                               power  = power, 
                                               sample.size.at.0.8 = sample.size.at.0.8)
              }else{
                dat.s.res[[s]] <- data.frame(Strain = s,
                                             t.stat = NA, 
                                             t.p    = NA, 
                                             lo.CI  = NA, 
                                             up.CI  = NA, 
                                             power  = NA, 
                                             sample.size.at.0.8 = NA)
              }
          }
              dat.s.res <- do.call(rbind.data.frame, dat.s.res)
              
              #merge y and dat.s.res
              ys <- full_join(y, dat.s.res, by = "Strain") %>%
                mutate(sig = ifelse(t.p <= 0.05, "*", "NS"),
                       StrainSig = paste0(Strain, "[", sig, ";", n, "]"))
              
              #plot based on SEM
              p1 <- ggplot(ys, aes(y = mean, x = reorder(StrainSig, -diff_mean), 
                                   color = sig, label = sig)) + 
                geom_point() + 
                geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.5, size = 1) +
                theme_cowplot() + 
                scale_color_hue(h = c(90, 180)) +
                geom_hline(yintercept = unique(ys$pop_mean)) +
                theme(axis.text.x = element_text(size = 16,angle = 90, hjust = 1), 
                      axis.text.y = element_text(size = 16),
                      legend.position="none") +
                scale_size_continuous(breaks = c(2,4,6,8,10,20), labels = c(2,4,6,8,10,20))+
                theme(text = element_text(size = 20)) +
                xlab("") +
                ylab(paste0(i, " (Mean ± SEM)")) +
                #ylim(min(ys$lo.CI, na.rm = TRUE), max(ys$up.CI, na.rm = TRUE)) +
                ggtitle(unique(ys$Sex)) +
                theme(axis.title.x = element_blank(),
                      plot.title = element_text(hjust = 0.5),
                      axis.text.x=element_text(angle = 90, hjust = 0.95, vjust = 0.5))
              #plot based on CI
              p2 <- ggplot(ys, aes(y = mean, x = reorder(StrainSig, -diff_mean), 
                                   color = sig, label = sig)) + 
                geom_point() + 
                geom_errorbar(aes(ymin = lo.CI, ymax = up.CI), width = 0.5, size = 1) +
                theme_cowplot() + 
                scale_color_hue(h = c(90, 180)) +
                geom_hline(yintercept = unique(ys$pop_mean)) +
                theme(axis.text.x = element_text(size = 16,angle = 90, hjust = 1), 
                      axis.text.y = element_text(size = 16),
                      legend.position="right") +
                scale_size_continuous(breaks = c(2,4,6,8,10,20), labels = c(2,4,6,8,10,20))+
                theme(text = element_text(size = 20)) +
                xlab("") +
                ylab(paste0(i, " (Mean 95% CI)")) +
                #ylim(min(ys$lo.CI, na.rm = TRUE), max(ys$up.CI, na.rm = TRUE)) +
                ggtitle(unique(ys$Sex)) +
                theme(axis.title.x = element_blank(),
                      plot.title = element_text(hjust = 0.5),
                      axis.text.x=element_text(angle = 90, hjust = 0.95, vjust = 0.5))
              return(list(ys, p1, p2))
        })
      }
  return(extremesCC)
}
```

### 4, Read data
```{r Read data, warning = FALSE, message = FALSE}
csv <- list.files(path = "data/pheno/Updated_CSNA_Data", pattern = ".csv", full.names = TRUE)
data <- list()
for(i in csv){
  print(i)
  data[[i]] <- readr::read_csv(i, col_names = TRUE, na = c("", "NA", "N/A", "NaN")) %>%
    rename_with(~str_remove(.x, "\\_cc|\\_do")) %>%
    rename_with(~(.x = "Strain"),   contains("_Strain")) %>%
    rename_with(~(.x = "Mouse.ID"), contains("Subject")) %>%
    rename_with(~(.x = "Sex"),      contains("Sex")) %>%
    mutate(Sex = case_when(
      Sex == "F" ~ "Female",
      Sex == "M" ~ "Male",
      TRUE       ~ as.character(Sex))) %>% #make Sex coding consistent
    mutate(across(Mouse.ID, as.character)) %>%
    mutate(Strain = case_when( #recode the founder strain to avoid inconsistency
      str_detect(Strain, "A/")    ~ "AJ",
      str_detect(Strain, "C57BL") ~ "B6",
      str_detect(Strain, "129S1") ~ "129",
      str_detect(Strain, "NOD")   ~ "NOD",
      str_detect(Strain, "NZO")   ~ "NZO",
      str_detect(Strain, "CAST")  ~ "CAST",
      str_detect(Strain, "PWK")   ~ "PWK",
      str_detect(Strain, "WSB")   ~ "WSB",
      TRUE                        ~ as.character(Strain)
    ))

  print("Column names : ");         print(colnames(data[[i]]))
  print("Strain names : ");         print(unique(data[[i]]$Strain))
  print("Duplicated Mouse.ID  : "); print(data[[i]]$Mouse.ID[duplicated(data[[i]]$Mouse.ID)])
}
names(data) <- csv

#RL_Acq-------------------------------------------------------------------------
RL_Acq <- data[which(str_detect(csv, "Acquisition"))] %>% 
  map(function(x){
    x %>% select(c("Mouse.ID", "Strain", "Sex", 
                   "Acq.Total.Trials", 
                   "Acq.Total.Correct", 
                   "Acq.Total.Omit", 
                   "Acq.Anticipatory.Correct.Responses", 
                   "Acq.Anticipatory.Incorrect.Responses")) %>%
      mutate(across(starts_with("Acq"), as.numeric)) %>%
      mutate(Acq.Anticipatory.Correct.Responses = case_when(
        Acq.Anticipatory.Correct.Responses >= 100 ~ NA_real_,
        is.na(Acq.Anticipatory.Correct.Responses) ~ NA_real_,
        TRUE                                      ~ Acq.Anticipatory.Correct.Responses)) %>% #outliers
      mutate(Acq.Anticipatory.Incorrect.Responses = case_when(
        Acq.Anticipatory.Incorrect.Responses >= 100 ~ NA_real_,
        is.na(Acq.Anticipatory.Incorrect.Responses) ~ NA_real_,
        TRUE                                        ~ Acq.Anticipatory.Incorrect.Responses)) %>%
      distinct(Mouse.ID, .keep_all = TRUE) #remove duplicates
}) %>%
  set_names(., nm = str_match(names(.),"2020_(\\w+?)\\.")[,2]) %>% .[order(names(.))]

#RL_Rev-------------------------------------------------------------------------
RL_Rev <- data[which(str_detect(csv, "Reversal"))] %>% 
  map(function(x){
    x %>% select(c("Mouse.ID", "Strain", "Sex", 
                   "Rev.Total.Trials", 
                   "Rev.Total.Correct", 
                   "Rev.Total.Omit", 
                   "Rev.Anticipatory.Correct.Responses", 
                   "Rev.Anticipatory.Incorrect.Responses")) %>%
      mutate(across(starts_with("Rev"), as.numeric)) %>%
      mutate(Rev.Anticipatory.Correct.Responses = case_when(
        Rev.Anticipatory.Correct.Responses >= 100 ~ NA_real_,
        is.na(Rev.Anticipatory.Correct.Responses) ~ NA_real_,
        TRUE                                      ~ Rev.Anticipatory.Correct.Responses)) %>% #outliers
      mutate(Rev.Anticipatory.Incorrect.Responses = case_when(
        Rev.Anticipatory.Incorrect.Responses >= 100 ~ NA_real_,
        is.na(Rev.Anticipatory.Incorrect.Responses) ~ NA_real_,
        TRUE                                        ~ Rev.Anticipatory.Incorrect.Responses)) %>%
      distinct(Mouse.ID, .keep_all = TRUE) #remove duplicates
}) %>%
  set_names(., nm = str_match(names(.),"2020_(\\w+?)\\.")[,2]) %>% .[order(names(.))]

#add Diff.Total.Trials
Diff <- map2(RL_Acq, RL_Rev, function(x,y){
  full_join(x, y, by = c("Mouse.ID", "Strain", "Sex")) %>%
    mutate(Diff.Total.Trials = Rev.Total.Trials - Acq.Total.Trials) %>% #Diff.Total.Trials
    select(c("Mouse.ID", "Strain", "Sex", "Diff.Total.Trials"))
})
#merge to RL_Rev
RL_Rev <- map2(RL_Rev, Diff, function(x,y){
    left_join(x, y, by = c("Mouse.ID", "Strain", "Sex"))
})

#SENS_COCA-------------------------------------------------------------------------
SENS_COCA <- data[which(str_detect(csv, "Sensitization"))] %>% 
  map(function(x){
    x %>% 
      filter(Study == "SENS-COCA") %>%
      select(c("Mouse.ID", "Strain", "Sex", 
               "D01_SAL",
               "D02_SAL",
               "D3_D2",
               "D5_D3",
               "D11_D3",
               "D12_D2",
               "D19_D11",
               "AUC")) %>%
      mutate(D1_D2 = D01_SAL - D02_SAL, .after = D02_SAL) %>%
      distinct(Mouse.ID, .keep_all = TRUE) #remove duplicates
}) %>%
  set_names(., nm = str_match(names(.),"2020_(\\w+?)\\.")[,2]) %>% .[order(names(.))]

#SENS-SHAM-------------------------------------------------------------------------
SENS_SHAM <- data[which(str_detect(csv, "Sensitization"))] %>% 
  map(function(x){
    x %>% 
      filter(Study == "SENS-SHAM") %>%
      select(c("Mouse.ID", "Strain", "Sex", 
               "D01_SAL",
               "D02_SAL",
               "D3_D2",
               "D5_D3",
               "D11_D3",
               "D12_D2",
               "D19_D11",
               "AUC")) %>%
      mutate(D1_D2 = D01_SAL - D02_SAL, .after = D02_SAL) %>%
      distinct(Mouse.ID, .keep_all = TRUE) #remove duplicates
}) %>% 
  set_names(., nm = str_match(names(.),"2020_(\\w+?)\\.")[,2]) %>% .[order(names(.))]

#IVSA-------------------------------------------------------------------------
IVSA <- data[which(str_detect(csv, "IVSA"))] %>% 
  map(function(x){
    x %>% 
      filter(!(DE_Exit_Reason == "FD")) %>%
      select(c("Mouse.ID", "Strain", "Sex",
               "AQ_SessionsToAcquisition",
               "DR_Inf_Total_1p0mgkg",
               "DR_Inf_Total_0p032mgkg",
               "DR_Inf_Total_0p1mgkg",
               "DR_Inf_Total_0p32mgkg",
               #"DR_Inf_Total_AUC",
               "EX_ALP_Total_s01",
               "EX_ILP_Total_s01",
               "EX_ALP_Total_s02",
               "EX_ILP_Total_s02",
               #"EX_Slope_ALP",
               #"EX_Slope_ILP",
               "RI_ALP_Total_s01",
               "RI_ILP_Total_s01",
               "RI_ALP_Total_s02",
               "RI_ILP_Total_s02",
               "RI_Inf_Total_s01",
               "RI_Inf_Total_s02")) %>%
      distinct(Mouse.ID, .keep_all = TRUE) #remove duplicates
}) %>%
  set_names(., nm = str_match(names(.),"2020_(\\w+?)\\.")[,2]) %>% .[order(names(.))]

#CR-------------------------------------------------------------------------
CR <- data[which(str_detect(csv, "CR"))] %>% 
  map(function(x){
    x %>% 
      select(c("Mouse.ID", "Strain", "Sex", 
               "period", 
               "phase",
               "amplitude",
               "damping",
               contains("of_fit"))) %>%
    rename_with(~(.x = "goodness_of_fit"), contains("of_fit"))
}) %>%
  set_names(., nm = str_match(names(.),"2020_(\\w+?)\\.")[,2]) %>% .[order(names(.))]

# note that Prj05_CR_preqc_12072020_DO.csv contains "J:DO", "CC075/UncJ", "CC030/GeniUncJ"
CR$DO <- CR$DO %>%
  filter(Strain == "J:DO")
#add CC from DO to CC for CR
CR$CC <- CR$DO %>%
  filter(Strain != "J:DO") %>%
  rbind(CR$CC)
```

### 5, Data summary
```{r Data summary, warning = FALSE}
#get result of summary
res.summary <- map(list(RL_Acq, RL_Rev, SENS_COCA, SENS_SHAM, IVSA, CR), function(x){
  print(names(x))
  y <- map(x, function(x){# for CC, DO and Founders
    x %>%
      group_by(Strain) %>% #Strain
      summarise(across(where(is.numeric), summary_func))
  })
  names(y) <- names(x)
  return(do.call(rbind.data.frame, y)) #combine CC, DO and Founders
}) %>%
  set_names(., nm = c("RL_Acq", "RL_Rev", "SENS_COCA", "SENS_SHAM", "IVSA", "CR"))
```

### 6, Plot summary
```{r Plot summary, warning = FALSE, fig.width=12, fig.height=12}
#pheno list in each domain
pheno.list <- list(RL_Acq, 
                   RL_Rev, 
                   SENS_COCA,
                   SENS_SHAM,
                   IVSA,
                   CR) %>% 
  map(~colnames(.x[[1]])[-c(1:3)]) %>% # get pheno names except for Mouse.ID, Strain, Sex
  set_names(., nm = c("RL_Acq", "RL_Rev", "SENS_COCA", "SENS_SHAM", "IVSA", "CR"))

#output location and name
out.pdf = as.list(paste0("output/pdf/mean.",names(pheno.list))) 
stopifnot(all.equal(names(pheno.list), names(res.summary)))

#plot for each domain
args1 <- list(summary.dat = res.summary, pheno.list = pheno.list, out.pdf = out.pdf)
args1 %>% 
  pmap(plot.summary)
```

### 7, Correlation in CC and_or Founders
```{r Correlation in CC and_or Founders, warning = FALSE, fig.width=16, fig.height=16}
#correlation in CC-------------------------
corr.cc <- res.summary[-4] %>% 
  reduce(full_join, by = "Strain") %>% # Don't include SENS_SHAM
  filter(str_detect(Strain, "^CC")) %>%
  select(c(contains("_mean"))) %>%
  rename_with(~str_remove(.x, "_mean"), contains("_mean")) %>%
  corr.test(use = "pairwise.complete.obs", method = "spearman", adjust = "fdr")
#plot p < 0.1
plot.corr.cc <- ggcorrplot(corr.cc$r, 
                           hc.order = FALSE, 
                           method = "square",
                           type = "lower",
                           outline.col = "grey", 
                           sig.level = 0.10, 
                           p.mat = corr.cc$p, 
                           insig="blank",
                           lab = TRUE,
                           lab_size = 3,
                           tl.cex = 18,
                           ggtheme = ggplot2::theme_bw,
                           title = "Cross Project Correlations in CC (p < 0.1)")
print(plot.corr.cc)
ggsave(file = paste0("output/pdf/corr.cc",".pdf"), 
         plot = plot.corr.cc,
         device = "pdf", height = 20, width = 20)
#no p cutoff
plot.corr.cc2 <- ggcorrplot(corr.cc$r, 
                           hc.order = FALSE, 
                           method = "square",
                           type = "lower",
                           outline.col = "grey", 
                           sig.level = 1, 
                           p.mat = corr.cc$p, 
                           insig="blank",
                           lab = TRUE,
                           lab_size = 3,
                           tl.cex = 18,
                           ggtheme = ggplot2::theme_bw,
                           title = "Cross Project Correlations in CC (no cutoff at p)")
print(plot.corr.cc2)
ggsave(file = paste0("output/pdf/corr.cc.fullmat",".pdf"), 
       plot = plot.corr.cc2,
       device = "pdf", height = 20, width = 20)
#save full matrix
save(corr.cc, file = "output/corr.cc.fullmat.RData")

#correlation in founders-------------------------
corr.fd <- res.summary[-4] %>% 
  reduce(full_join, by = "Strain") %>% # Don't include SENS_SHAM
  filter(!str_detect(Strain, "^CC")) %>%
  filter(Strain != "J:DO") %>%
  select(c(contains("_mean"))) %>%
  rename_with(~str_remove(.x, "_mean"), contains("_mean")) %>%
  corr.test(use = "pairwise.complete.obs", method = "spearman", adjust = "fdr")
#plot p < 0.1
plot.corr.fd <- ggcorrplot(corr.fd$r, 
                           hc.order = FALSE, 
                           method = "square",
                           type = "lower",
                           outline.col = "grey", 
                           sig.level = 0.10, 
                           p.mat = corr.fd$p, 
                           insig="blank",
                           lab = TRUE,
                           lab_size = 3,
                           tl.cex = 18,
                           ggtheme = ggplot2::theme_bw,
                           title = "Cross Project Correlations in Founders (p < 0.1)")
print(plot.corr.fd)
ggsave(file = paste0("output/pdf/corr.fd",".pdf"), 
         plot = plot.corr.fd,
         device = "pdf", height = 20, width = 20)
#no p cutoff
plot.corr.fd2 <- ggcorrplot(corr.fd$r, 
                           hc.order = FALSE, 
                           method = "square",
                           type = "lower",
                           outline.col = "grey", 
                           sig.level = 1, 
                           p.mat = corr.fd$p, 
                           insig="blank",
                           lab = TRUE,
                           lab_size = 3,
                           tl.cex = 18,
                           ggtheme = ggplot2::theme_bw,
                           title = "Cross Project Correlations in Founders (no cutoff at p)")
print(plot.corr.fd2)
ggsave(file = paste0("output/pdf/corr.fd.fullmat",".pdf"), 
       plot = plot.corr.fd2,
       device = "pdf", height = 20, width = 20)
#save full matrix
save(corr.fd, file = "output/corr.fd.fullmat.RData")

#correlation in cc + founders-------------------------
corr.cc.fd <- res.summary[-4] %>% 
  reduce(full_join, by = "Strain") %>% # Don't include SENS_SHAM
  filter(Strain != "J:DO") %>%
  select(c(contains("_mean"))) %>%
  rename_with(~str_remove(.x, "_mean"), contains("_mean")) %>%
  corr.test(use = "pairwise.complete.obs", method = "spearman", adjust = "fdr")
#plot p < 0.1
plot.corr.cc.fd <- ggcorrplot(corr.cc.fd$r, 
                           hc.order = FALSE, 
                           method = "square",
                           type = "lower",
                           outline.col = "grey", 
                           sig.level = 0.10, 
                           p.mat = corr.cc.fd$p, 
                           insig="blank",
                           lab = TRUE,
                           lab_size = 3,
                           tl.cex = 18,
                           ggtheme = ggplot2::theme_bw,
                           title = "Cross Project Correlations in CC and Founders (p < 0.1)")
print(plot.corr.cc.fd)
ggsave(file = paste0("output/pdf/corr.cc.fd",".pdf"), 
       plot = plot.corr.cc.fd,
       device = "pdf", height = 20, width = 20)
#no p cutoff
plot.corr.cc.fd2 <- ggcorrplot(corr.cc.fd$r, 
                              hc.order = FALSE, 
                              method = "square",
                              type = "lower",
                              outline.col = "grey", 
                              sig.level = 1, 
                              p.mat = corr.cc.fd$p, 
                              insig="blank",
                              lab = TRUE,
                              lab_size = 3,
                              tl.cex = 18,
                              ggtheme = ggplot2::theme_bw,
                              title = "Cross Project Correlations in CC and Founders (no cutoff at p)")
print(plot.corr.cc.fd2)
ggsave(file = paste0("output/pdf/corr.cc.fd.fullmat",".pdf"), 
       plot = plot.corr.cc.fd2,
       device = "pdf", height = 20, width = 20)
#save full matrix
save(corr.cc.fd, file = "output/corr.cc.fd.fullmat.RData")
```

### 8, CC Extreme Strain Survey
```{r, CC Extreme Strain Survey, warning = FALSE, fig.width=24, fig.height=12}
#CC Extreme Strain Survey
args2 <- list(cc.data = list(RL_Acq    = RL_Acq$CC, 
                             RL_Rev    = RL_Rev$CC, 
                             SENS_COCA = SENS_COCA$CC, 
                             IVSA      = IVSA$CC, 
                             CR        = CR$CC), #CC data across projects
              pheno.list = pheno.list[c(-4)]) #no SENS_SHAM
extremesCC.res <- args2 %>% 
  pmap(extremesCC_func)

#print a pdf for each domain
for(d in names(extremesCC.res)){
  extremesCC.res[[d]] %>% 
    modify_depth(2, ~print(plot_grid(.x[[2]], .x[[3]]))) #print in rmarkdown
  #print into pdf
  pdf(file = paste0("output/pdf/", d, ".extremesCC.pdf"), width = 24, height = 10)
    extremesCC.res[[d]] %>% modify_depth(2, ~print(plot_grid(.x[[2]], .x[[3]])))
  dev.off()
}
```

### 9, DO phenos and correlation
```{r DO phenos and correlation, warning = FALSE, eval = TRUE}
# Novelty data process
# This will process CSNA phenotype in
# OFA_raw_07012020.csv
# NPP_raw_07012020.csv
# LD_raw_07012020.csv
# HB_raw_07012020.csv.
# 1, remove outliers
# 2, get residuals
# 3, rankz transformation on residuals

if(TRUE){# it's done in the cluster
  #Process OFA phenotype data  -----------------------------------------------------------------------
  #mouse id < 10000 are pre-center, anything > 1000 is post-center
  #raw OFA
  ofa.raw <- read.csv("data/pheno/OFA_raw_07012020.csv", header = TRUE)
  ofa.raw$Birth.Date <- as.Date(ofa.raw$Birth.Date, "%m/%d/%Y")
  ofa.raw$date.of.test <- as.Date(ofa.raw$date.of.test, "%m/%d/%y")
  ofa.raw <- as_tibble(ofa.raw)

  # pre center
  ofa.pre.nooutlier <- ofa.raw %>%
    filter(Mouse.ID < 10000) %>%
    group_by(Sex) %>%
    mutate(across(total.distance.traveled:distance.traveled.last.five,remove_outliers))

  #regression
  ofa.pre.resids <- ofa.pre.nooutlier
  #OFA
  ofa.pre.resids[,20:43] <- apply(ofa.pre.nooutlier[,20:43], 2, FUN = function(x, data = ofa.pre.nooutlier){
    data$date.of.test <- as.factor(data$date.of.test)
    data$tester <- as.factor(data$tester)
    m <- lmer(x ~ tester + (1|date.of.test), data = data, na.action = na.exclude, REML = TRUE)
    r <- residuals(m)
    r
  })

  #rankz transform
  ofa.pre.rankz.resids <- ofa.pre.resids
  ofa.pre.rankz.resids[,20:43] <- apply(ofa.pre.resids[,20:43], 2, rz.transform)

  # post center
  ofa.post.nooutlier <- ofa.raw %>%
    filter(Mouse.ID >= 10000) %>%
    group_by(Sex) %>%
    mutate(across(total.distance.traveled:distance.traveled.last.five,remove_outliers))

  #regression
  ofa.post.resids <- ofa.post.nooutlier
  #OFA
  ofa.post.resids[,20:43] <- apply(ofa.post.nooutlier[,20:43], 2, FUN = function(x, data = ofa.post.nooutlier){
    data$date.of.test <- as.factor(data$date.of.test)
    data$tester <- as.factor(data$tester)
    m <- lmer(x ~ tester + (1|date.of.test), data = data, na.action = na.exclude, REML = TRUE)
    r <- residuals(m)
    r
  })

  #rankz transform
  ofa.post.rankz.resids <- ofa.post.resids
  ofa.post.rankz.resids[,20:43] <- apply(ofa.post.resids[,20:43], 2, rz.transform)

  #write out the csv
  colnames(ofa.raw)[20:43] <- paste0("OFA.", colnames(ofa.raw)[20:43],             ".raw")
  ofa.nooutlier <- rbind.data.frame(ofa.pre.nooutlier, ofa.post.nooutlier)
  colnames(ofa.nooutlier)[20:43] <- paste0("OFA.", colnames(ofa.nooutlier)[20:43], ".nooutlier")
  #write.csv(ofa.nooutlier, file = "data/pheno/OFA_nooutlier_07012020.csv", row.names = F, quote = F)
  ofa.resids <- rbind.data.frame(ofa.pre.resids, ofa.post.resids)
  colnames(ofa.resids)[20:43] <- paste0("OFA.", colnames(ofa.resids)[20:43],       ".resids")
  #write.csv(ofa.resids, file = "data/pheno/OFA_resids_07012020.csv", row.names = F, quote = F)
  ofa.ranknorm <- rbind.data.frame(ofa.pre.rankz.resids, ofa.post.rankz.resids)
  colnames(ofa.ranknorm)[20:43] <- paste0("OFA.", colnames(ofa.ranknorm)[20:43],   ".ranknorm")
  #write.csv(ofa.ranknorm, file = "data/pheno/OFA_resids_ranknorm_07012020.csv", row.names = F, quote = F)

  #mege ofa
  OFA <- list(ofa.raw,
              ofa.nooutlier,
              ofa.resids,
              ofa.ranknorm) %>%
    reduce(full_join, by = colnames(ofa.raw)[1:19])
  colnames(OFA)[2:19] <- paste0("OFA.", colnames(OFA)[2:19])

  #Process NPP phenotype data  -----------------------------------------------------------------------
  #raw npp
  npp.raw <- read.csv("data/pheno/NPP_raw_07012020.csv", header = TRUE)
  npp.raw$Birth.Date <- as.Date(npp.raw$Birth.Date, "%m/%d/%Y")
  npp.raw <- npp.raw[,!(colnames(npp.raw)  %in% c("Test.MovementCounts_GreyZone_Total","Exposure.MovementCounts_GreyZone_Total"))]
  npp.raw$date.of.test <- as.Date(npp.raw$date.of.test, "%m/%d/%y")
  npp.raw <- as_tibble(npp.raw)

  # pre center
  npp.pre.nooutlier <- npp.raw %>%
    filter(Mouse.ID < 10000) %>%
    group_by(Sex) %>%
    mutate(across(Exposure.ExplorationCounts_BlackZone_Total:NoveltyPreference.MovementCounts_GreyWhiteBlack_Total,remove_outliers))

  # post center
  npp.post.nooutlier <- npp.raw %>%
    filter(Mouse.ID >= 10000) %>%
    group_by(Sex) %>%
    mutate(across(Exposure.ExplorationCounts_BlackZone_Total:NoveltyPreference.MovementCounts_GreyWhiteBlack_Total,remove_outliers))

  npp.nooutlier <- rbind.data.frame(npp.pre.nooutlier, npp.post.nooutlier)
  #regression
  npp.resids <- npp.nooutlier
  #npp
  npp.resids[,41:78] <- apply(npp.nooutlier[,41:78], 2, FUN = function(x, data = npp.nooutlier){
    data$date.of.test <- as.factor(data$date.of.test)
    data$tester <- as.factor(data$tester)
    m <- lmer(x ~ tester + (1|date.of.test), data = data, na.action = na.exclude, REML = TRUE)
    r <- residuals(m)
    r
  })

  #rankz transform
  npp.ranknorm <- npp.resids
  npp.ranknorm[,41:78] <- apply(npp.resids[,41:78], 2, rz.transform)

  #write out the csv
  colnames(npp.raw)[41:78]       <- paste0("NPP.", colnames(npp.raw)[41:78],       ".raw")
  colnames(npp.nooutlier)[41:78] <- paste0("NPP.", colnames(npp.nooutlier)[41:78], ".nooutlier")
  colnames(npp.resids)[41:78]    <- paste0("NPP.", colnames(npp.resids)[41:78],    ".resids")
  colnames(npp.ranknorm)[41:78]  <- paste0("NPP.", colnames(npp.ranknorm)[41:78],  ".ranknorm")
  #write.csv(npp.nooutlier, file = "data/pheno/NPP_nooutlier_07012020.csv", row.names = F, quote = F)
  #write.csv(npp.resids, file = "data/pheno/NPP_resids_07012020.csv", row.names = F, quote = F)
  #write.csv(npp.ranknorm, file = "data/pheno/NPP_resids_ranknorm_07012020.csv", row.names = F, quote = F)

  #mege NPP
  NPP <- list(npp.raw,
              npp.nooutlier,
              npp.resids,
              npp.ranknorm) %>%
    reduce(full_join, by = colnames(npp.raw)[-c(41:78)])
  colnames(NPP)[c(2:40,79,80)] <- paste0("NPP.", colnames(NPP)[c(2:40,79,80)])

  #Process LD phenotype data -----------------------------------------------------------------------
  #raw LD
  LD.raw <- read.csv("data/pheno/LD_raw_07012020.csv", header = TRUE)
  LD.raw$Birth.Date <- as.Date(LD.raw$Birth.Date, "%m/%d/%Y")
  LD.raw$date.of.test <- as.Date(LD.raw$date.of.test, "%m/%d/%y")
  LD.raw <- as_tibble(LD.raw)

  # pre center
  LD.pre.nooutlier <- LD.raw %>%
    filter(Mouse.ID < 10000) %>%
    group_by(Sex) %>%
    mutate(across(pct.distance.traveled.in.light:total.transitions,remove_outliers))

  #regression
  LD.pre.resids <- LD.pre.nooutlier
  #LD
  LD.pre.resids[,20:25] <- apply(LD.pre.nooutlier[,20:25], 2, FUN = function(x, data = LD.pre.nooutlier){
    data$date.of.test <- as.factor(data$date.of.test)
    m <- lm(x ~ date.of.test, data = data, na.action = na.exclude)
    r <- residuals(m)
    r
  })

  #rankz transform
  LD.pre.rankz.resids <- LD.pre.resids
  LD.pre.rankz.resids[,20:25] <- apply(LD.pre.resids[,20:25], 2, rz.transform)

  # post center
  LD.post.nooutlier <- LD.raw %>%
    filter(Mouse.ID >= 10000) %>%
    group_by(Sex) %>%
    mutate(across(pct.distance.traveled.in.light:total.transitions,remove_outliers))

  #regression
  LD.post.resids <- LD.post.nooutlier
  #LD
  LD.post.resids[,20:25] <- apply(LD.post.nooutlier[,20:25], 2, FUN = function(x, data = LD.post.nooutlier){
    data$date.of.test <- as.factor(data$date.of.test)
    data$tester <- as.factor(data$tester)
    m <- lmer(x ~ tester + (1|date.of.test), data = data, na.action = na.exclude, REML = TRUE)
    r <- residuals(m)
    r
  })

  #rankz transform
  LD.post.rankz.resids <- LD.post.resids
  LD.post.rankz.resids[,20:25] <- apply(LD.post.resids[,20:25], 2, rz.transform)

  #write out the csv
  colnames(LD.raw)[20:25]       <- paste0("LD.", colnames(LD.raw)[20:25],       ".raw")
  LD.nooutlier <- rbind.data.frame(LD.pre.nooutlier, LD.post.nooutlier)
  colnames(LD.nooutlier)[20:25] <- paste0("LD.", colnames(LD.nooutlier)[20:25], ".nooutlier")
  #write.csv(LD.nooutlier, file = "data/pheno/LD_nooutlier_07012020.csv", row.names = F, quote = F)
  LD.resids <- rbind.data.frame(LD.pre.resids, LD.post.resids)
  colnames(LD.resids)[20:25]    <- paste0("LD.", colnames(LD.resids)[20:25],    ".resids")
  #write.csv(LD.resids, file = "data/pheno/LD_resids_07012020.csv", row.names = F, quote = F)
  LD.ranknorm <- rbind.data.frame(LD.pre.rankz.resids, LD.post.rankz.resids)
  colnames(LD.ranknorm)[20:25]  <- paste0("LD.", colnames(LD.ranknorm)[20:25],  ".ranknorm")
  #write.csv(LD.ranknorm, file = "data/pheno/LD_resids_ranknorm_07012020.csv", row.names = F, quote = F)

  #mege LD
  LD <- list(LD.raw,
             LD.nooutlier,
             LD.resids,
             LD.ranknorm) %>%
    reduce(full_join, by = colnames(LD.raw)[1:19])
  colnames(LD)[c(2:19)] <- paste0("LD.", colnames(LD)[c(2:19)])

  #Process HB phenotype data -----------------------------------------------------------------------
  #raw HB
  HB.raw <- read.csv("data/pheno/HB_raw_07012020.csv", header = TRUE)
  HB.raw$Birth.Date <- as.Date(HB.raw$Birth.Date, "%m/%d/%y")
  HB.raw$date.of.test <- as.Date(HB.raw$date.of.test, "%m/%d/%y")
  HB.raw <- as_tibble(HB.raw)

  # pre center
  HB.pre.nooutlier <- HB.raw %>%
    filter(Mouse.ID < 10000) %>%
    group_by(Sex) %>%
    mutate(across(Total.Entries:Repeat.Entries,remove_outliers))

  #regression
  HB.pre.resids <- HB.pre.nooutlier
  #HB
  HB.pre.resids[,20:22] <- apply(HB.pre.nooutlier[,20:22], 2, FUN = function(x, data = HB.pre.nooutlier){
    data$date.of.test <- as.factor(data$date.of.test)
    m <- lm(x ~ date.of.test, data = data, na.action = na.exclude)
    r <- residuals(m)
    r
  })

  #rankz transform
  HB.pre.rankz.resids <- HB.pre.resids
  HB.pre.rankz.resids[,20:22] <- apply(HB.pre.resids[,20:22], 2, rz.transform)

  # post center
  HB.post.nooutlier <- HB.raw %>%
    filter(Mouse.ID >= 10000) %>%
    group_by(Sex) %>%
    mutate(across(Total.Entries:Repeat.Entries,remove_outliers))

  #regression
  HB.post.resids <- HB.post.nooutlier
  #HB
  HB.post.resids[,20:22] <- apply(HB.post.nooutlier[,20:22], 2, FUN = function(x, data = HB.post.nooutlier){
    data$date.of.test <- as.factor(data$date.of.test)
    data$tester <- as.factor(data$tester)
    m <- lmer(x ~ tester + (1|date.of.test), data = data, na.action = na.exclude, REML = TRUE)
    r <- residuals(m)
    r
  })

  #rankz transform
  HB.post.rankz.resids <- HB.post.resids
  HB.post.rankz.resids[,20:22] <- apply(HB.post.resids[,20:22], 2, rz.transform)

  #write out the csv
  colnames(HB.raw)[20:22]       <- paste0("HB.", colnames(HB.raw)[20:22],       ".raw")
  HB.nooutlier <- rbind.data.frame(HB.pre.nooutlier, HB.post.nooutlier)
  colnames(HB.nooutlier)[20:22] <- paste0("HB.", colnames(HB.nooutlier)[20:22], ".nooutlier")
  #write.csv(HB.nooutlier, file = "data/pheno/HB_nooutlier_07012020.csv", row.names = F, quote = F)
  HB.resids <- rbind.data.frame(HB.pre.resids, HB.post.resids)
  colnames(HB.resids)[20:22]    <- paste0("HB.", colnames(HB.resids)[20:22],    ".resids")
  #write.csv(HB.resids, file = "data/pheno/HB_resids_07012020.csv", row.names = F, quote = F)
  HB.ranknorm <- rbind.data.frame(HB.pre.rankz.resids, HB.post.rankz.resids)
  colnames(HB.ranknorm)[20:22]  <- paste0("HB.", colnames(HB.ranknorm)[20:22],  ".ranknorm")
  #write.csv(HB.ranknorm, file = "data/pheno/HB_resids_ranknorm_07012020.csv", row.names = F, quote = F)

  #mege HB
  HB <- list(HB.raw,
             HB.nooutlier,
             HB.resids,
             HB.ranknorm) %>%
    reduce(full_join, by = colnames(HB.raw)[1:19])
  colnames(HB)[c(2:19)] <- paste0("HB.", colnames(HB)[c(2:19)])
}

# Read each project
#sens----------------------------------------------------------
sens <- read.csv("data/pheno/Prj02_SENS_pheno_qttl2_DO_12012020_COCA_all.csv")
sens <- sens %>% 
  rename(Mouse.ID = mouseID) %>%
  mutate(across(Mouse.ID, as.character)) %>%
  left_join(data$`data/pheno/Updated_CSNA_Data/Prj02_Sensitization_preqc_12282020_DO.csv`[,c(1,8,9)], by = "Mouse.ID") %>%
  distinct(Mouse.ID, .keep_all = TRUE) %>%
  mutate(
    D1_D2 = D01_SAL - D02_SAL,
    D3_D2.nooutlier = case_when(
      (D3_D2 < -15000 | D3_D2 > 50000) ~ NA_real_,
      TRUE ~ D3_D2),
    D5_D3.nooutlier = case_when(
      (D5_D3 < -25000) ~ NA_real_,
      TRUE ~ D5_D3),
    D11_D3.nooutlier = case_when(
      (D11_D3 > 50000 | D11_D3 < -25000) ~ NA_real_,
      TRUE ~ D11_D3),
    D12_D2.nooutlier = case_when(
      (D12_D2 > 20000 | D12_D2 < -20000) ~ NA_real_,
      TRUE ~ D12_D2),
    D19_D11.nooutlier = case_when(
      (D19_D11 > 20000 | D19_D11 < -20000) ~ NA_real_,
      TRUE ~ D19_D11),
    AUC.nooutlier = case_when(
      (AUC > 50000 | AUC < -20000) ~ NA_real_,
      TRUE ~ AUC)
  )#remove outliers
colnames(sens)[4:12] <- paste0(colnames(sens)[4:12], ".raw")
colnames(sens)[4:18] <- paste0("SENS.", colnames(sens)[4:18])

#cr----------------------------------------------------------
cr <- read.csv("data/pheno/Prj05_rythms_do.mean.csv")
colnames(cr) <- c("ID", "Period.raw", "Phase.raw", "Amplitude.raw", "Damping.raw", "Goodness.of.fit.raw", "Mouse.ID")
cr <- cr %>% filter(!(ID %in% c("26046_5", "20662_8"))) %>% # remove duplicates
  mutate(across(Mouse.ID, as.character)) %>%
  mutate(Damping.nooutlier = case_when(
    (Damping.raw >= 20 | Damping.raw <= -2) ~ NA_real_,
    TRUE ~ Damping.raw
  ))
colnames(cr)[c(2:6,8)] <- paste0("CR.", colnames(cr)[c(2:6,8)])

#rl----------------------------------------------------------
rl <- readr::read_csv("data/pheno/Prj01_RL_pheno_qttl2_DO_11092020.csv", na = c("N/A", "NA"))
rl <- rl %>% rename(Mouse.ID = mouseID) %>% distinct(Mouse.ID, .keep_all = TRUE) %>%
  mutate(across(Mouse.ID, as.character))
colnames(rl)[4:8] <- paste0(colnames(rl)[4:8], ".raw")
rl <- rl %>% mutate(
  Acq.Anticipatory.Correct.Responses.nooutlier = case_when(
    (Acq.Anticipatory.Correct.Responses.raw > 5 | is.na(Acq.Anticipatory.Correct.Responses.raw)) ~ NA_real_,
    TRUE ~ Acq.Anticipatory.Correct.Responses.raw),
  Rev.Anticipatory.Incorrect.Responses.nooutlier = case_when(
    (Rev.Anticipatory.Incorrect.Responses.raw >6) ~ NA_real_,
    TRUE ~ Rev.Anticipatory.Incorrect.Responses.raw
  ))#omit values >6 for Acq.Anticipatory.Correct.Responses and Rev.Anticipatory.Incorrect.Responses variables.
colnames(rl)[4:10] <- paste0("RL.", colnames(rl)[4:10])

#ivsa----------------------------------------------------------
ivsa <- IVSA$DO %>%
  rename_with(~paste0("IVSA.", .x, ".raw"), contains("_"))

#merge OFA, NPP, LD, HB, sens, cr, rl, ivsa for DO----------------------
DO.allpheno <- list(OFA, NPP, LD, HB, sens, cr, rl, ivsa) %>%
  map(~(.x %>% mutate(across(Mouse.ID, as.character)))) %>%
  reduce(full_join, by = "Mouse.ID")
#correlation on raw phenotypes
DO.allrawpheno.corr <- DO.allpheno %>%
  select(contains(".raw")) %>%
  corr.test(use = "pairwise.complete.obs", 
            method = "spearman", 
            adjust = "fdr")
#plot
plot.DO.allrawpheno.corr <- ggcorrplot(DO.allrawpheno.corr$r, 
                                       hc.order = FALSE, 
                                       method = "square",
                                       type = "lower",
                                       outline.col = "grey", 
                                       sig.level = 0.10, 
                                       p.mat = DO.allrawpheno.corr$p, 
                                       insig="blank",
                                       lab = TRUE,
                                       lab_size = 3,
                                       tl.cex = 16,
                                       ggtheme = ggplot2::theme_bw,
                                       title = "Cross Project Correlations in DO (p < 0.1)")
#save full matrix
save(DO.allrawpheno.corr, file = "output/DO.allrawpheno.corr.fullmat.RData")

# print(plot.DO.allpheno.corr)
# ggsave(file = paste0("output/pdf/DO.allpheno.corr",".pdf"), 
#          plot = plot.DO.allpheno.corr,
#          device = "pdf", height = 20, width = 20)
# NEED phenotype list for plotting
```

### 10, Phenotypes for QTL
```{r Phenotypes for QTL, warning = FALSE, eval = TRUE}
#merge with genotype data after QC----------------------------------------------------------
load("data/Jackson_Lab_12_batches/gm_DO3173_qc_newid.RData")
csna.allqtl.pheno <- gm_after_qc$covar %>%
  rename(Mouse.ID = name) %>%
  left_join(DO.allpheno, by = "Mouse.ID") %>%
  select(c(colnames(.)[1:5], matches("raw|outlier|resids|ranknorm")))
write.csv(csna.allqtl.pheno, file  = "data/pheno/csna.allqtl.pheno.csv", row.names = F)
```
