uthor: "Hao He"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Genotype diagnostics for diversity outbred mice
We first load the R/qtl2 package and the data.
We’ll also load the R/broman package for some utilities and plotting functions, and R/qtlcharts for interactive graphs.

# library
```{r library, message=FALSE}
library(broman)
library(qtl2)
library(qtlcharts)
library(ggplot2)
library(ggrepel)
library(DOQTL)
library(mclust)
source("/projects/csna/csna_workflow/code/reconst_utils.R")
```

# Missing data per sample
```{r Missing data per sample,fig.height=12, fig.width=12}
load("/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/Gagnon_gm.RData")
gm

percent_missing <- n_missing(gm, "ind", "prop")*100
setScreenSize(height=100, width=300)
labels <- paste0(names(percent_missing), " (", round(percent_missing,2), "%)")
iplot(seq_along(percent_missing), percent_missing, indID=labels,
      chartOpts=list(xlab="Mouse", ylab="Percent missing genotype data",
                     ylim=c(0, 60)))

#save into pdf
pdf(file = "/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/Percent_missing_genotype_data.pdf", width = 20, height = 20)
labels <- as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "V01_"))[,2])
labels[percent_missing < 5] = ""
# Change point shapes and colors
p <- ggplot(data = data.frame(Mouse=seq_along(percent_missing),  
                         Percent_missing_genotype_data = percent_missing,
                         batch = factor(as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,5]))), 
        aes(x=Mouse, y=Percent_missing_genotype_data, color = batch)) +
  geom_point() +
  geom_hline(yintercept=5, linetype="solid", color = "red") +
  geom_text_repel(aes(label=labels), vjust = 0, nudge_y = 0.01, show.legend = FALSE, size=3) +
  theme(text = element_text(size = 20))
p
dev.off()

p

save(percent_missing,
     file = "/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/percent_missing_id.RData")
```

# Sexes
```{r Sexes}
xint <- read_csv_numer("/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/Jackson_Lab_Gagnon_qtl2_chrXint.csv", transpose=TRUE)
yint <- read_csv_numer("/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/Jackson_Lab_Gagnon_qtl2_chrYint.csv", transpose=TRUE)

# Gigamuga marker annotation file from UNC.
gm_marker_file = "http://csbio.unc.edu/MUGA/snps.gigamuga.Rdata" #FIXED
# Read in the UNC GigaMUGA SNPs and clusters.
load(url(gm_marker_file))
#subset down to gm
snps$marker = as.character(snps$marker)

#load the intensities.fst.RData
load("/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/intensities.fst.RData")
#X and Y channel
X <- result[result$channel == "X",]
rownames(X) <- X$snp
X <- X[,c(-1,-2)]

Y <- result[result$channel == "Y",]
rownames(Y) <- Y$snp
Y <- Y[,c(-1,-2)]

#determine sex
sex = determine_sex_chry_m(x = X, y = Y, markers = snps)$sex

gm$covar <- merge(data.frame(id = names(sex),
                             predict.sex = sex,stringsAsFactors = F),
                   gm$covar,
                   by.x = "id", 
                   by.y = "row.names")
rownames(gm$covar) <- gm$covar$id

#sex order
sex <- gm$covar[rownames(xint),"predict.sex"]

x_pval <- apply(xint, 2, function(a) t.test(a ~ sex)$p.value)
y_pval <- apply(yint, 2, function(a) t.test(a ~ sex)$p.value)

xint_ave <- rowMeans(xint[, x_pval < 0.05/length(x_pval)], na.rm=TRUE)
yint_ave <- rowMeans(yint[, y_pval < 0.05/length(y_pval)], na.rm=TRUE)

point_colors <- as.character( brocolors("web")[c("green", "purple")] )
labels <- paste0(names(xint_ave))
iplot(xint_ave, yint_ave, group=sex, indID=labels,
      chartOpts=list(pointcolor=point_colors, pointsize=4,
                     xlab="Average X chr intensity", ylab="Average Y chr intensity"))

phetX <- rowSums(gm$geno$X == 2)/rowSums(gm$geno$X != 0)
phetX <- phetX[names(phetX) %in% names(xint_ave)]
iplot(xint_ave, phetX, group=sex, indID=labels,
      chartOpts=list(pointcolor=point_colors, pointsize=4,
                     xlab="Average X chr intensity", ylab="Proportion het on X chr"))
```

# Sample duplicates
```{r Sample duplicates}
cg <- compare_geno(gm, cores=10)
summary.cg <- summary(cg)
summary.cg
summary.cg$Name.ind1 <- as.character(do.call(rbind.data.frame, strsplit(as.character(summary.cg$ind1), "_"))[,6])
summary.cg$Name.ind2 <- as.character(do.call(rbind.data.frame, strsplit(as.character(summary.cg$ind2), "_"))[,6])
summary.cg$miss.ind1 <- percent_missing[match(summary.cg$ind1, names(percent_missing))]
summary.cg$miss.ind2 <- percent_missing[match(summary.cg$ind2, names(percent_missing))]
summary.cg$remove.id <- ifelse(summary.cg$miss.ind1 > summary.cg$miss.ind2, summary.cg$ind1, summary.cg$ind2)
summary.cg$remove.id  

save(summary.cg,
     file = "/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/summary.cg.RData")

pdf(file = "/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/Proportion_matching_genotypes_before_removal_of_bad_samples.pdf", width = 20, height = 20) 
par(mar=c(5.1,0.6,0.6, 0.6))
hist(cg[upper.tri(cg)], breaks=seq(0, 1, length=201),
     main="", yaxt="n", ylab="", xlab="Proportion matching genotypes")
rug(cg[upper.tri(cg)])
dev.off()

par(mar=c(5.1,0.6,0.6, 0.6))
hist(cg[upper.tri(cg)], breaks=seq(0, 1, length=201),
     main="", yaxt="n", ylab="", xlab="Proportion matching genotypes")
rug(cg[upper.tri(cg)])

pdf(file = "/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/Proportion_matching_genotypes_after_removal_of_bad_samples.pdf",width = 20, height = 20) 
cgsub <- cg[percent_missing < 5, percent_missing < 5]
par(mar=c(5.1,0.6,0.6, 0.6))
hist(cgsub[upper.tri(cgsub)], breaks=seq(0, 1, length=201),
     main="", yaxt="n", ylab="", xlab="Proportion matching genotypes")
rug(cgsub[upper.tri(cgsub)])
dev.off()

cgsub <- cg[percent_missing < 5, percent_missing < 5]
par(mar=c(5.1,0.6,0.6, 0.6))
hist(cgsub[upper.tri(cgsub)], breaks=seq(0, 1, length=201),
     main="", yaxt="n", ylab="", xlab="Proportion matching genotypes")
rug(cgsub[upper.tri(cgsub)])

#show top 20 samples with missing genotypes
percent_missing <- n_missing(gm, "ind", "prop")*100
round(sort(percent_missing, decreasing=TRUE)[1:20], 1)
```

# Array intensities and Genotype frequencies
```{r Array intensities and Genotype frequencies, fig.height=12, fig.width=12}
int <- result
rm(result)
int <- int[seq(1, nrow(int), by=2),-(1:2)] + int[-seq(1, nrow(int), by=2),-(1:2)]
int <- int[,intersect(ind_ids(gm), colnames(int))]
n <- names(sort(percent_missing[intersect(ind_ids(gm), colnames(int))], decreasing=TRUE))
iboxplot(log10(t(int[,n])+1), orderByMedian=FALSE, chartOpts=list(ylab="log10(SNP intensity + 1)"))

# Genotype frequencies
g <- do.call("cbind", gm$geno[1:19])
fg <- do.call("cbind", gm$founder_geno[1:19])
g <- g[,colSums(fg==0)==0]
fg <- fg[,colSums(fg==0)==0]
fgn <- colSums(fg==3)

gf_ind <- vector("list", 4)
for(i in 1:4) {
  gf_ind[[i]] <- t(apply(g[,fgn==i], 1, function(a) table(factor(a, 1:3))/sum(a != 0)))
}

par(mfrow=c(4,1), mar=c(0.6, 0.6, 2.6, 0.6))
for(i in 1:4) {
  triplot(c("AA", "AB", "BB"), main=paste0("MAF = ", i, "/8"))
  tripoints(gf_ind[[i]], pch=21, bg="lightblue")
  tripoints(c((1-i/8)^2, 2*i/8*(1-i/8), (i/8)^2), pch=21, bg="violetred")
  
  if(i>=3) { # label mouse with lowest het
    wh <- which(gf_ind[[i]][,2] == min(gf_ind[[i]][,2]))
    tritext(gf_ind[[i]][wh,,drop=FALSE] + c(0.02, -0.02, 0),
            names(wh), adj=c(0, 1))
  }
  
  # label other mice
  if(i==1) {
    lab <- rownames(gf_ind[[i]])[gf_ind[[i]][,2]>0.3]
  }
  else if(i==2) {
    lab <- rownames(gf_ind[[i]])[gf_ind[[i]][,2]>0.48]
  }
  else if(i==3) {
    lab <- rownames(gf_ind[[i]])[gf_ind[[i]][,2]>0.51]
  }
  else if(i==4) {
    lab <- rownames(gf_ind[[i]])[gf_ind[[i]][,2]>0.6]
  }
  
  for(ind in lab) {
    if(grepl("^F", ind) && i != 3) {
      tritext(gf_ind[[i]][ind,,drop=FALSE] + c(-0.01, 0, +0.01), ind, adj=c(1,0.5))
    } else {
      tritext(gf_ind[[i]][ind,,drop=FALSE] + c(0.01, 0, -0.01), ind, adj=c(0,0.5))
    }
  }
}

```

# Crossover counts and Genotyping error LOD scores
```{r Crossover counts and Genotyping error LOD scores}
# qsub qsub pr_cross_over.sh to get the pr, m and nxo
#load pre-caluated results
load("/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/pr.RData")
load("/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/m.RData")
load("/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/nxo.RData")

#crossover
totxo <- rowSums(nxo)[rownames(gm$covar)]
iplot(seq_along(totxo),
      totxo,
      group=gm$covar$ngen,
      chartOpts=list(xlab="Mouse", ylab="Number of crossovers", 
                     margin=list(left=80,top=40,right=40,bottom=40,inner=5),
                     axispos=list(xtitle=25,ytitle=50,xlabel=5,ylabel=5)))

#save crossover into pdf
pdf(file = "/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/number_crossover.pdf")
cross_over <- data.frame(Mouse = seq_along(totxo), Number_crossovers = totxo, generation = gm$covar$ngen)
names(totxo) <- as.character(do.call(rbind.data.frame, strsplit(names(totxo), "V01_"))[,2])
names(totxo)[totxo <= 800] = ""
# Change point shapes and colors
p <-ggplot(cross_over, aes(x=Mouse, y=Number_crossovers, fill = generation, color=generation)) +
  geom_point() +
  geom_text_repel(aes(label=names(totxo),hjust=0,vjust=0), show.legend = FALSE)
p
dev.off()

p

#Here are the crossover counts for those  mice:
tmp <- cbind(percent_missing=round(percent_missing,2), total_xo=totxo)[percent_missing >= 5,]
tmp[order(tmp[,1]),]

# Genotyping error LOD scores
load("/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/e.RData")
errors_ind <- rowSums(e>2)/n_typed(gm)*100
lab <- paste0(names(errors_ind), " (", myround(percent_missing,1), "%)")
iplot(seq_along(errors_ind), errors_ind, indID=lab,
      chartOpts=list(xlab="Mouse", ylab="Percent genotyping errors", ylim=c(0, 8),
                     axispos=list(xtitle=25, ytitle=50, xlabel=5, ylabel=5)))
save(errors_ind, file = "/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/errors_ind.RData")

# Apparent genotyping errors
load("/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/snpg.RData")

gobs <- do.call("cbind", gm$geno)
gobs[gobs==0] <- NA

par(pty="s")
err_direct <- rowMeans(snpg != gobs, na.rm=TRUE)*100
errors_ind_0 <- rowSums(e > 0)/n_typed(gm)*100
par(mar=c(4.1,4.1,0.6, 0.6))
grayplot(errors_ind_0, err_direct,
         xlab="Percent errors (error LOD > 0)",
         ylab="Percent errors (obs vs predicted)",
         xlim=c(0, 2), ylim=c(0, 2))
abline(0,1,lty=2, col="gray60")

pdf(file = "/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/Percent_genotype_errors_obs_vs_predicted.pdf",width = 20, height = 20) 
par(pty="s")
err_direct <- rowMeans(snpg != gobs, na.rm=TRUE)*100
errors_ind_0 <- rowSums(e > 0)/n_typed(gm)*100
par(mar=c(4.1,4.1,0.6, 0.6))
grayplot(errors_ind_0, err_direct,
         xlab="Percent errors (error LOD > 0)",
         ylab="Percent errors (obs vs predicted)",
         xlim=c(0, 2), ylim=c(0, 2))
abline(0,1,lty=2, col="gray60")
dev.off()
```

# Missing data in Markers and Genotype frequencies Markers
```{r Missing data in Markers and Genotype frequencies Markers, eval=TRUE}
#It can also be useful to look at the proportion of missing genotypes by marker. 
#Markers with a lot of missing data were likely difficult to call, and so the genotypes that were called may contain a lot of errors.
pmis_mar <- n_missing(gm, "marker", "proportion")*100

par(mar=c(5.1,0.6,0.6, 0.6))
hist(pmis_mar, breaks=seq(0, 100, length=201),
     main="", yaxt="n", ylab="", xlab="Percent missing genotypes")
rug(pmis_mar)

pdf(file = "/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/Percent_missing_genotype_data_per_marker.pdf")
par(mar=c(5.1,0.6,0.6, 0.6))
hist(pmis_mar, breaks=seq(0, 100, length=201),
     main="", yaxt="n", ylab="", xlab="Percent missing genotypes")
rug(pmis_mar)
dev.off()

# Genotype frequencies Markers
gf_mar <- t(apply(g, 2, function(a) table(factor(a, 1:3))/sum(a != 0)))
gn_mar <- t(apply(g, 2, function(a) table(factor(a, 1:3))))

pdf(file = "/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/genotype_frequency_marker.pdf")
par(mfrow=c(2,2), mar=c(0.6, 0.6, 2.6, 0.6))
for(i in 1:4) {
  triplot(c("AA", "AB", "BB"), main=paste0("MAF = ", i, "/8"))
  z <- gf_mar[fgn==i,]
  z <- z[rowSums(is.na(z)) < 3,]
  tripoints(z, pch=21, bg="gray80", cex=0.6)
  tripoints(c((1-i/8)^2, 2*i/8*(1-i/8), (i/8)^2), pch=21, bg="violetred")
}
dev.off()

par(mfrow=c(2,2), mar=c(0.6, 0.6, 2.6, 0.6))
for(i in 1:4) {
  triplot(c("AA", "AB", "BB"), main=paste0("MAF = ", i, "/8"))
  z <- gf_mar[fgn==i,]
  z <- z[rowSums(is.na(z)) < 3,]
  tripoints(z, pch=21, bg="gray80", cex=0.6)
  tripoints(c((1-i/8)^2, 2*i/8*(1-i/8), (i/8)^2), pch=21, bg="violetred")
}

# Genotype errors Markers
errors_mar <- colSums(e>2)/n_typed(gm, "marker")*100

grayplot(pmis_mar, errors_mar,
         xlab="Proportion missing", ylab="Proportion genotyping errors")

pdf(file = "/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/genotype_error_marker.pdf")
grayplot(pmis_mar, errors_mar,
         xlab="Proportion missing", ylab="Proportion genotyping errors")
dev.off()
```

# Remove bad samples
```{r remove bad samples, eval=F}
#percent missing
qc_info <- merge(gm$covar,
                 data.frame(id = names(percent_missing),
                            percent_missing = percent_missing,stringsAsFactors = F),by = "id")
#cross_over
qc_info <- merge(qc_info,
                 data.frame(id = rownames(cross_over),
                            Number_crossovers = cross_over$Number_crossovers,stringsAsFactors = F),by = "id")

#missing sex
qc_info[qc_info$id %in% tot.id.covar.csna.sexna$ID, "sex"] <- NA
qc_info$sex.match <- ifelse(qc_info$predict.sex == qc_info$sex, TRUE, FALSE)

#genotype errors
qc_info <- merge(qc_info,
                 data.frame(id = names(errors_ind),
                            genotype_erros = errors_ind,stringsAsFactors = F),by = "id")
#duplicated id to be remove
qc_info$remove.id.duplicated <- ifelse(qc_info$id %in% summary.cg$remove.id, TRUE,FALSE)

bad.sample <- qc_info[qc_info$ngen ==1 | qc_info$Number_crossovers <= 200 | qc_info$Number_crossovers >=1000 | qc_info$percent_missing >= 10 | qc_info$genotype_erros >= 1 | qc_info$remove.id.duplicated == TRUE,]

save(qc_info, bad.sample, file = "/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/qc_info_bad_sample.RData")

#remove 77 bad samples
gm.no.bad <- gm[paste0("-",as.character(bad.sample$id)),]

#fix sex mismatch
gm.no.bad$covar[gm.no.bad$covar$id %in% names(yint_ave[!names(yint_ave) %in% bad.sample$id][yint_ave[!names(yint_ave) %in% bad.sample$id] <= 0.1]),"predict.sex"] <- "F"

gm.no.bad

#2437 subjects
# update other stuff
e <- e[ind_ids(gm.no.bad),]
g <- g[ind_ids(gm.no.bad),]
snpg <- snpg[ind_ids(gm.no.bad),]

length(errors_mar[errors_mar > 5])
# omit the 273 markers with error rates >5%.
bad_markers <- find_markerpos(gm.no.bad, names(errors_mar[errors_mar > 5]))
save(bad_markers, file = "/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/bad_markers.RData")
#drop bad markers
gm_DO2437_qc <- drop_markers(gm.no.bad, names(errors_mar)[errors_mar > 5])

gm_DO2437_qc

save(gm_DO2437_qc, file = "/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/gm_DO2437_qc.RData")
save(e,g,snpg, file = "/projects/csna/csna_workflow/data/Jackson_Lab_Gagnon/e_g_snpg_qc.RData")
```
